---
title: "analysis-correlations-amplitude"
author: "Abel Serrano Juste"
date: "`r Sys.Date()`"
output: html_document
---

# INITIALIZATION
SET GLOBAL VARS
```{r}
PATH = dirname(rstudioapi::getSourceEditorContext()$path)
print(PATH)
setwd(PATH)

PLACE = 'Corbalan'
```


Load common libraries and data set
```{r}
source('init-analysis.R')
```

Import other libraries
```{r}
library(ggnewscale)
```

# Calculate amplitude for all trees by class

```{r}
# First, let's create an amplitude.df to store the output.
column_names <- c("date", "min", "max", "ampl", "class", "series")
amplitude.indiv = data.frame(matrix(nrow = 0, ncol = length(column_names)))
colnames(amplitude.indiv) <- column_names
for (dendro.no in unique(db$series)) {

  # Filter data by that no
  dat = db[db$series == dendro.no,]
  class = first(dat$class)
  dat = dat %>% select(ts, value)
  
  # Max - min daily amplitude
  dat.ampl <- calculate_amplitude(dat)
  aux <- cbind(ampl = dat.ampl, series = as.factor(dendro.no), class = class) %>% rename (date = ampl.date, min = ampl.min, max = ampl.max, ampl = ampl.ampl)
  amplitude.indiv = rbind.data.frame(amplitude.indiv, aux)
}
```

Create one different dataframe for every class
```{r}
amplitude.Qi <- amplitude.indiv[amplitude.indiv$class == "Quercus",]
amplitude.P_ND <- amplitude.indiv[amplitude.indiv$class == "ND",]
amplitude.P_D <- amplitude.indiv[amplitude.indiv$class == "D",]
```


## VISUALIZATION

Plot amplitudes:
```{r}
# Define color scales for each class
green_colors <- c("lightgreen", "green", "darkgreen", "darkolivegreen", "olivedrab", "springgreen")
blue_colors <- c("blue", "cyan", "aquamarine", "cadetblue", "dodgerblue", "darkblue")
orange_colors <- c("orange", "orangered", "tomato", "darkorange", "coral", "tan")

ggplot() +
  ggtitle("Amplitudes painted by color scale according to its group") +
  geom_line(data = amplitude.P_D, aes(x = date, y = ampl, col = series)) +
  scale_color_manual(values = orange_colors, name = "Declining trees") +
  new_scale_color() +
  geom_line(data = amplitude.P_ND, aes(x = date, y = ampl, col = series), ) +
  scale_color_manual(values = blue_colors, name = "Non-Declining trees") +
  new_scale_color() +
  geom_line(data = amplitude.Qi, aes(x = date, y = ampl, col = series)) +
  scale_color_manual(values = green_colors, name = "Quercus Ilex") +
  labs(x=expression(''),
       y=expression(Delta*" D (um)")) +
  scale_x_date(date_breaks = "1 month", date_labels="%b %Y") +
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

```{r}
boxplot(ampl ~ class, data = amplitude.indiv)
```


```{r}
data <- amplitude.indiv %>% filter(series == first(unique(amplitude.P_D$series)) | series == first(unique(amplitude.Qi$series)) | series == first(unique(amplitude.P_ND$series)))
ggplot() +
  ggtitle("Amplitudes painted by color scale according to its group") +
  geom_line(data = data, aes(x = date, y = ampl, col = series)) +
  scale_x_date(date_breaks = "1 month", date_labels="%b %Y")
```

Plot histograms:
```{r}
for (dendro.no in unique(amplitude.indiv$series)) {
  dat = amplitude.indiv %>% filter(series == dendro.no)
  hist(log(dat$ampl))
}
```

# Analysis

## ANOVA to explore significant difference between means by class
Analysis ANOVA to see if there's significant difference between the average amplitude of each class

```{r}
aov.model <- aov(ampl ~ class + series, data = amplitude.indiv)
summary(aov.model)
```

```{r}
plot(aov.model$residuals)
```


```{r}
TukeyHSD(aov.model)
```
```{r}
plot(TukeyHSD(aov.model))
```


## Correlations

### Aggregate data

With all daily amplitudes, calculate mean of each one by class

```{r}
amplitude.all = amplitude.indiv %>% group_by(date) %>% summarise(mean = mean(ampl), .groups = "drop")
amplitude.all
amplitude.df = amplitude.indiv %>% group_by(date, class) %>% summarise(mean = mean(ampl))
amplitude.df
amplitude.Qi <- amplitude.df[amplitude.df$class == "Quercus",]
amplitude.P_ND <- amplitude.df[amplitude.df$class == "ND",]
amplitude.P_D <- amplitude.df[amplitude.df$class == "D",]
```

### Amplitude ~ climate

Here we show a plot with VWC vs amplitude per class

```{r}
ggplot(data = amplitude.df, mapping = aes(x = date, y = mean, col = class)) +
  ggtitle(glue('Soil moisture vs daily stem amplitude difference for {PLACE}')) +
  geom_line() +
  scale_y_continuous("Amplitude (um)", breaks=seq(0,100,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = range.vwc * 100, linetype = "Max - Min Volumetric Water Content (%)"), col="blue")
```

Following, we will explore correlation within microclimate variables (Soil Moisture and temperature) and daily seasonal amplitude.

## Amplitude ~ Temperature
Explore correlation within temperature and daily stem amplitude difference (Not significant)

```{r}
cor.test(
  #specify the two variables to correlate
  amplitude.P_D$mean, clim.daily$interquartil.temp,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
find_Max_CCF(amplitude.P_D$mean, clim.daily$interquartil.temp)
```

## Amplitude ~ Soil Moisture
Explore cross-correlation within Soil Moisture and daily seasonal amplitude

### for all trees and classes:

```{r}
ccf (amplitude.all$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
sort_CCF_values(amplitude.all$mean, clim.daily$max.vwc)
```

### for Quercus Ilex:

Plot VWC and daily amplitude for Qi:

```{r}
plot(clim.daily$max.vwc, amplitude.Qi$mean)
```

```{r}
plot(log10(clim.daily$max.vwc), log10(amplitude.Qi$mean))
```

Test correlation:
```{r}
cor.test(
  #specify the two variables to correlate
  clim.daily$max.vwc, amplitude.Qi$mean,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
hist(log10(amplitude.Qi$mean+1))
```
```{r}
hist(log10(clim.daily$max.vwc+1))
```

Standard correlations using different methods:
```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


Cross-correlations:
```{r}
ccf (amplitude.Qi$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.Qi$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
find_Max_CCF(amplitude.Qi$mean, clim.daily$max.vwc)
sort_CCF_values(amplitude.Qi$mean, clim.daily$max.vwc)
```

### For P_ND: amplitude ~ VWC 

Exploring plots VWC ~ daily amplitude:

```{r}
plot(clim.daily$max.vwc, amplitude.P_ND$mean)
# plot(lm(log10(clim.daily$mean.vwc) ~ log10(amplitude.P_ND$mean)))
```


```{r}
plot(log10(clim.daily$max.vwc), log10(amplitude.P_ND$mean))
# plot(lm(clim.daily$mean.vwc ~ amplitude.P_ND$mean))
```


```{r}
join.df <- full_join(clim.daily, amplitude.P_ND, by="date")
  ggplot(data = join.df, aes(log10(max.vwc), log10(mean))) +
  # black points graph
  geom_point() +
  # add correlation with errors and blue color
  stat_smooth(method = 'lm',
              method.args = list(start= c(a = 1,b=1)),
              se=T, color = "blue") +
   # theme
  theme_classic() +
  # add labels
  labs( x = "Max Soil Moisture per day - Log",
        y = "amplitude variations (um) - Log")
```
Exploring correlations:
```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


```{r}
ccf (amplitude.P_ND$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
sort_CCF_values(amplitude.P_ND$mean, clim.daily$max.vwc)
sort_CCF_values(log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc))
```

## With P_D

```{r}
cor.test(log10(amplitude.P_D$mean), log10(clim.daily$max.vwc),
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_D$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_D$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


```{r}
ccf (amplitude.P_D$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.P_D$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
sort_CCF_values(amplitude.P_D$mean, clim.daily$max.vwc)
sort_CCF_values(log10(amplitude.P_D$mean), log10(clim.daily$max.vwc))
```

