---
title: "TS-analysis"
author: "Abel Serrano Juste"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages:
```{r}
source('lib-dendro.R')
library(glue)
library(tidyverse)
library (ggplot2)
library(lubridate)
library(plotly)

library(imputeTS)
library(anomalize) # Decomposition
library(stats) # Decomposition
library(ggpubr) # ggarange
```

Set global vars
```{r}
### DEFINE GLOBAL VARS ###
PATH = dirname(rstudioapi::getSourceEditorContext()$path)
print(PATH)
setwd(PATH)

PLACE = 'Miedes'
DATA_DIR = 'processed/Miedes-last'
ENV_DIR = 'processed/Miedes-env-processed'
```

Select dendros for this analysis
```{r}
selected_dendros <- c("92222156", "92222169", "92222175", "92222157", "92222154", "92222170", "92222173", "92222180", "92222155", "92222163", "92222171", "92222161", "92222164")
```


Load dataset:
```{r}
list_files <- list.files(file.path(".",DATA_DIR), pattern="*.csv$", full.names=TRUE)
db<-read.all.processed(list_files)
# db <- read.one.processed(file.path(".",DATA_DIR, 'proc-M_Qi1-92222155.csv'))
```
## CLEAN & PREPARE DATA ###

keep data of selected dendros only
```{r}
db = db %>% filter(db$series %in% selected_dendros)
str(db)
```

select only growing season of 2023
```{r}
# Set initial and final date for analysis
ts_start <- "2023-03-15 09:00:00" # 2 days after installation
ts_end <-"2023-09-10 07:00:00" # last timestamp of downloaded data
db <- reset.initial.values(db, ts_start, ts_end)
```



# INSPECT DATA
```{r}
str(db)
head(db)
tail(db)
```

# Data imputation

Missing data
```{r}
statsNA(db$value)
```

Let's fill it through interpolation
```{r}
db$value <- db$value %>% 
            na_interpolation(option = "spline")
```

```{r}
# db_locf$ts = as_datetime(db_locf$ts, tz = 'Europe/Madrid')
```

```{r}
# plot_ly(db, x = ~ts, y = ~value, color=~series, type = 'scatter', mode = 'lines')
```

# Time Series decomposition

```{r}
## Function decompose
plot_decompose <- function (db, class) {
  # decompose a time series (has to be in tibble format)
  decompose = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select varaible to decompose
    value)
  
  # plot time series
  p_ts = ggplot (decompose, aes(x = ts, y = observed)) +
    ggtitle(glue("Time series decomposition for {class}")) +
    geom_line (col = "black") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = NULL,
             y = "Original data (um)")
  
  # plot trend
  p_trend = ggplot (decompose, aes(x = ts, y = trend)) +
    geom_line (col = "#D55E00") + 
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    geom_hline(yintercept = 0, linetype='dotted', col = 'red') +
    theme_classic() +
    labs ( x = NULL,
             y = "Trend")
  
  # plot season
  p_season = ggplot (decompose, aes(x = ts, y = season)) +
    geom_line (col = "#E69F00") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = NULL,
             y = "Season")
  
  # plot remainder
  p_remainder = ggplot (decompose, aes(x = ts, y = remainder)) +
    geom_line (col = "#F0E442") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = "Period (month)",
             y = "Remainder")
  
  # plot all together
  ggarrange (p_ts, p_trend, p_season, p_remainder, ncol = 1)
}
```


# by individuals


First individual: 92222154 - Non-Declining Pinus Pinater
```{r}
dendro.no = 92222154
dat = db[db$series == dendro.no,]
ts_start <- "2023-05-01 09:00:00" # From first of May
ts_end <-"2023-09-10 07:00:00" # last timestamp of downloaded data
dat = dat %>% select(ts, value)
dat = dat %>% filter(ts>=ts_start & ts<=ts_end)
dat.ts <- ts(data = dat$value,
             # start = dat$ts[1],
             # end = dat$ts[length(dat$ts)],
             frequency = 96)
head(dat.ts, n = 180)
```

## Basic ts decomposition
time series basic decomposition (constant seasonality)

```{r}
#png('output/decomp-anomalize.png', width=15, height=10, units="in", res=300)
plot_decompose(dat, dendro.no)
#dev.off()
#ggsave(glue('output/decomp-anomalize.png'), width = 15, height = 10)
```

## Daily amplitude
Daily amplitude, i.e., the difference between the highest and the lowest value.

```{r}
dat.ampl <- dat %>% mutate(date = date(ts)) %>% group_by(date) %>% summarize(max = max(value), min = min(value)) %>% mutate(ampl = max-min)
ggplot(data = dat.ampl, mapping = aes(x = date, y = ampl)) +
       geom_line()
```

## Using STL method

```{r}
stl.out = stl(dat.ts, s.window = 25, t.window = 673)
summary(stl.out)
plot(stl.out)
```

```{r}
png(glue('output/decomp-stl-{dendro.no}.png'), width=15, height=10, units="in", res=300)
plot(stl.out)
dev.off()
```

Now plotting only seasonality but with dates in the x-axis
```{r}
library(zoo)
seasonality <- stl.out$time.series[,1]
timestamps <- seq(from = as.POSIXct(ts_start, tz='Madrid/Spain'), by = "15 min", length.out = length(dat.ts))
zoo_data <- zoo(seasonality, order.by = timestamps)
plot(zoo_data, xaxt = "n", type = "l", xlab = "", ylab = "Value", main = "Time Series by Month-Year")
axis(1, at = time(zoo_data), labels = format(time(zoo_data), "%Y-%m"))

# Add x-axis label
mtext("Month-Year", side = 1, line = 3)
```



## Other methods
Testing other methods of decomposing seasonality:

```{r}
# library("seasonal")
# m <- seas(dat.ts)
# summary(m)
```

```{r}
library("stlplus")

dendro_stl <- stlplus(dat.ts, n.p = 96,
  s.window = 25, t.window = 673)
  #sub.labels = substr(month.name, 1, 3))

plot(dendro_stl, ylab = "stem growth (um)", xlab = "Time (days)")
plot_seasonal(dendro_stl)
plot_trend(dendro_stl)
plot_cycle(dendro_stl)
plot_rembycycle(dendro_stl)

```

Library forecast (automated stl)

```{r}
library(forecast)
dat.ts %>%
  mstl(iterate = 8) %>%
  autoplot()
```

Second individual: 92222155 - Quercus Ilex

Third individual: 92222169 - Declining Pinus Pinater


# aggregated decomposition

# Mean of trends
```{r}
d <- db %>% filter (series %in% P_D)
s = unique(d$series)

ts = d %>%  group_by(series) %>% pull(ts)
acum = vector(mode = "numeric", length = length(ts))

for (series_i in s) {
  dat = d[d$series == series_i,]
  
  print(head(dat))
  print(anyNA(dat))
  
  # decompose a time series (has to be in tibble format)
  decompose = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(dat,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select variable to decompose
    value)
  
    acum = acum + decompose$trend;
}

n = length(s)

mean = modify(acum, (\(x) x / n));
mean_of_trends = tibble(mean, ts)
```


```{r}
  # plot trend
p_trend = ggplot (mean_of_trends, aes(x = ts, y = mean)) +
  geom_line (col = "#D55E00") + 
  scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
  geom_hline(yintercept = 0, linetype='dotted', col = 'red') +
  theme_classic() +
  labs ( x = NULL,
           y = "Trend")
p_trend
```

## Mean by class
```{r}
TreeList<-read.table("TreeList.txt",header=T)

Qi = TreeList %>% filter(class == "Quercus") %>% pull(series)
P_D = TreeList %>% filter(class == "D") %>% pull(series)
P_ND = TreeList %>% filter(class == "ND") %>% pull(series)

db.Qi.agg <- db %>% filter (series %in% Qi) %>% group_by(ts) %>% dplyr::summarise(mean = mean(value, na.rm = T), sd = sd(value, na.rm = T) ) %>% rename(value = mean)
db.D.agg <- db %>% filter (series %in% P_D) %>% group_by(ts) %>% dplyr::summarise(mean = mean(value, na.rm = T), sd = sd(value, na.rm = T) ) %>% rename(value = mean)
db.ND.agg <- db %>% filter (series %in% P_ND) %>% group_by(ts) %>% dplyr::summarise(mean = mean(value, na.rm = T), sd = sd(value, na.rm = T) ) %>% rename(value = mean)
```


```{r}
plot_decompose(db.Qi.agg, "Quercus Ilex")
ggsave(glue('output/decomp-Qi-{PLACE}.png'), width = 15, height = 10)
```



```{r}
plot_decompose(db.D.agg, "Declining Pinus Pinaster")
ggsave(glue('output/decomp-D-{PLACE}.png'), width = 15, height = 10)
```


```{r}
plot_decompose(db.ND.agg, "Non-Declining Pinus Pinaster")
ggsave(glue('output/decomp-ND-{PLACE}.png'), width = 15, height = 10)
```

# Plot altogether

```{r}
## Function decompose
  decompose.Qi = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db.Qi.agg,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select variable to decompose
    value)
  
  decompose.D = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db.D.agg,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select variable to decompose
    value)
  
  decompose.ND = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db.ND.agg,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select variable to decompose
    value)
```


```{r}
 p_trend = ggplot ( ) +
  ggtitle('Trends by class') +
    geom_line (data = decompose.Qi, aes(x = ts, y = trend, col = "Quercus Ilex")) +
    geom_ribbon(data = db.Qi.agg, aes(x = ts, ymin=sd, ymax=sd), fill='lightgreen', alpha=0.3, show.legend	= FALSE, linetype = 0) +
    geom_line (data = decompose.D, aes(x = ts, y = trend, col = "Declining Pinus")) +
    geom_line (data = decompose.ND, aes(x = ts, y = trend, col = "Non-Declining Pinus")) +
    geom_hline(yintercept = 0, linetype='dotted', col = 'red') +
    # scale_color_manual(values = c("darkgreen", "aquamarine", "#E69F00")) +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = NULL,
           y = "Trend (um)",
           col = "Class")
p_trend
ggsave(glue('output/trends-byclass-{PLACE}.png'), width = 15, height = 10)
```


```{r}
decompose = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select varaible to decompose
    value)
```

```{r}
decompose = rbind.data.frame(mutate(decompose.Qi, class=factor("Quercus")), mutate(decompose.D, class=factor("D")), mutate(decompose.ND, class=factor("ND")))
```

All trend data with plotly
```{r}
plot_ly(decompose, x = ~ts, y = ~trend, color=~class, type = 'scatter', mode = 'lines')
```

# Exploring daily cycles in different periods

All seasonal data
```{r}
plot_ly(decompose, x = ~ts, y = ~season, color=~class, type = 'scatter', mode = 'lines')
```

Set different portion of data. i.e. outside the growing season.

```{r}
# Set initial and final date for analysis
ts_start <- "2023-08-28 09:00:00" # 2 days after installation
ts_end <-"2023-09-10 9:00:00" # last timestamp of downloaded data
db.2 <- reset.initial.values(db, ts_start, ts_end)

db.Qi.agg <- db.2 %>% filter (series %in% Qi) %>% group_by(ts) %>% dplyr::summarise(value = mean(value, na.rm = T), sd = sd(value, na.rm = T) )
db.D.agg <- db.2 %>% filter (series %in% P_D) %>% group_by(ts) %>% dplyr::summarise(value = mean(value, na.rm = T), sd = sd(value, na.rm = T) )
db.ND.agg <- db.2 %>% filter (series %in% P_ND) %>% group_by(ts) %>% dplyr::summarise(value = mean(value, na.rm = T), sd = sd(value, na.rm = T) )

## Function decompose
  decompose.Qi = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db.Qi.agg,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select variable to decompose
    value)
  
  decompose.D = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db.D.agg,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select variable to decompose
    value)
  
  decompose.ND = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db.ND.agg,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select variable to decompose
    value)
```


```{r}
decompose = rbind.data.frame(mutate(decompose.Qi, class=factor("Quercus")), mutate(decompose.D, class=factor("D")), mutate(decompose.ND, class=factor("ND")))
```

```{r}
plot_ly(decompose, x = ~ts, y = ~season, color=~class, type = 'scatter', mode = 'lines') %>%  layout(title="Growth II")
```
