---
title: "Seasonality-analysis"
author: "Abel Serrano Juste"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages:
```{r}
source('lib-dendro.R')
source('lib-ts-analysis.R')
library(glue)
library(tidyverse)
library(ggplot2)
theme_set(theme_bw())
library(lubridate)

library(imputeTS)
library(biwavelet)
```

Set global vars
```{r}
### DEFINE GLOBAL VARS ###
PATH = dirname(rstudioapi::getSourceEditorContext()$path)
print(PATH)
setwd(PATH)

PLACE = 'Corbalan'
DATA_DIR = 'processed/Corbalan-processed'
ENV_DIR = 'processed/Corbalan-env-processed'

TreeList <- read.table("TreeList.txt", header=T)
```

Select dendros for this analysis (all if FALSE)
```{r}
# selected_dendros <- c("92222156", "92222169", "92222175", "92222157", "92222154", "92222170", "92222173", "92222180", "92222155", "92222163", "92222171", "92222161", "92222164")
selected_dendros <- FALSE
```

Select TMS serial no for this analysis
```{r}
SELECTED_TMS <- "94231943"
```


Load dataset:
```{r}
list_files <- list.files(file.path(".",DATA_DIR), pattern="*.csv$", full.names=TRUE)
db<-read.all.processed(list_files)
```

## CLEAN & PREPARE DATA ###

keep data of selected dendros only
```{r}
if (selected_dendros) {db = db %>% filter(db$series %in% selected_dendros)}
str(db)
```

select from May 2023 to Sept 2023
```{r}
# Set initial and final date for analysis
ts_start <- "2023-05-01 09:00:00" # from first of May
ts_end <-"2023-09-10 07:00:00" # to 10th of september
db <- reset.initial.values(db, ts_start, ts_end)
```

# INSPECT DATA
```{r}
str(db)
head(db)
tail(db)
```

# Data imputation

Missing data
```{r}
statsNA(db$value)
```

Let's fill it through interpolation
```{r}
db$value <- db$value %>% 
            na_interpolation(option = "spline")
```

# Time Series decomposition

# Calculate mean Seasonality
Mean seasonality for all trees:
```{r}
seasonalities <- calculate_stl_seasonalities(db, unique(db$series))
summary(seasonalities)

seasonalities.agg <- seasonalities %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.agg)
```

```{r}
ggplot( data = seasonalities.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for all trees")
#ggsave(glue('output/{PLACE}-mean-seasonalities-all.png'), width = 15, height = 10)
```

# Calculate Seasonality by class

Calculate mean of the seasonality for Qi, PD and P_ND
```{r}
Qi = TreeList %>% filter(class == "Quercus", series %in% unique(db$series)) %>% pull(series)
P_D = TreeList %>% filter(class == "D", series %in% unique(db$series)) %>% pull(series)
P_ND = TreeList %>% filter(class == "ND", series %in% unique(db$series)) %>% pull(series)
```

Mean seasonality for Quercus Ilex:
```{r}
seasonalities.qi <- calculate_stl_seasonalities(db, Qi)
summary(seasonalities.qi)

seasonalities.qi.agg <- seasonalities.qi %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.qi.agg)
```

```{r}
ggplot( data = seasonalities.qi, mapping = aes(x=ts, y = value, color = series)) + 
  geom_line(alpha = 0.5) +
  theme_bw()
```


```{r}
ggplot( data = seasonalities.qi.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Quercus Ilex")
#ggsave(glue('output/{PLACE}-mean-seasonality-Qi.png'), width = 15, height = 10)
```

Mean seasonality for Non-Declining Pinus
```{r}
seasonalities.P_ND <- calculate_stl_seasonalities(db, P_ND)
summary(seasonalities.P_ND)

seasonalities.P_ND.agg <- seasonalities.P_ND %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.P_ND.agg)
```

```{r}
ggplot( data = seasonalities.P_ND, mapping = aes(x=ts, y = value, color = series)) + 
  geom_line(alpha = 0.5) +
  theme_bw()
```


```{r}
ggplot( data = seasonalities.P_ND.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Non-Declining Pinus")
#ggsave(glue('output/{PLACE}-mean-seasonality-P_ND.png'), width = 15, height = 10)
```

Mean seasonality for Declining Pinus
```{r}
seasonalities.P_D <- calculate_stl_seasonalities(db, P_D)
summary(seasonalities.P_D)

seasonalities.P_D.agg <- seasonalities.P_D %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.P_D.agg)
```

```{r}
ggplot( data = seasonalities.P_D, mapping = aes(x=ts, y = value, color = series)) + 
  geom_line(alpha = 0.5) +
  theme_bw()
```
```{r}
# library(plotly)
# plot_ly(
#   data = seasonalities.P_D,
#   x = ~ts,
#   y = ~value,
#   type = "scatter",
#   mode = "lines",
#   color = ~series
#   )
```



```{r}
ggplot( data = seasonalities.P_D.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Declining Pinus")
#ggsave(glue('output/{PLACE}-mean-seasonality-P_D.png'), width = 15, height = 10)
```

# Calculate amplitude for all trees by class

```{r}
# First, let's create an amplitude.df to store the output.
column_names <- c("date", "min", "max", "ampl", "class")
amplitude.df = data.frame(matrix(nrow = 0, ncol = length(column_names)))
colnames(amplitude.df) <- column_names
for (dendro.no in unique(db$series)) {
  
  # skip if we don't have the dendro registered in TreeList
  if (! (dendro.no %in% TreeList$series)) next;
  
  class = (TreeList[TreeList$series == dendro.no,])$class
  name = paste(dendro.no, class, sep="-")
  
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  
  # Max - min daily amplitude
  dat.ampl <- calculate_amplitude(dat)
  aux <- cbind(ampl = dat.ampl, name = name, class=class) %>% rename (date = ampl.date, min = ampl.min, max = ampl.max, ampl = ampl.ampl)
  amplitude.df = rbind.data.frame(amplitude.df, aux)
}
```

With all daily amplitudes, calculate mean of each one by class

```{r}
amplitude.all = amplitude.df %>% group_by(date) %>% summarise(mean = mean(ampl), .groups = "drop")
amplitude.all
amplitude.df = amplitude.df %>% group_by(date, class) %>% summarise(mean = mean(ampl))
amplitude.df
```

Create one different dataframe for every class
```{r}
amplitude.Qi <- amplitude.df[amplitude.df$class == "Quercus",]
amplitude.P_ND <- amplitude.df[amplitude.df$class == "ND",]
amplitude.P_D <- amplitude.df[amplitude.df$class == "D",]
```

# Climate data

Importing environmental data and keeping the one sensor with all valid data
```{r}
db.env <- read.env.proc(file.path(PATH,ENV_DIR,'proc-env.csv'))
db.env <- db.env[db.env$series == SELECTED_TMS,]
```

```{r}
#library(plotly)
#plot_ly(db.env, x = ~ts, y = ~vwc, color = ~series, type = 'scatter', mode = 'lines')
```


Filter env data to period of study
```{r}
db.env <- db.env[which(db.env$ts>=ts_start & db.env$ts<=ts_end),]
summary(db.env)
```

Plot env data:
```{r}
plot_temp_and_humidity <-
  ggplot(data = db.env, aes(x=ts, y=surface.temp)) +
    ggtitle(glue("Temperature and humidity for {PLACE} (May to September 2023)")) +
    geom_line(aes(colour = surface.temp)) +
    theme_bw() +
    scale_colour_gradient(low = "light blue", high = "red") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme(axis.text.x = element_text(angle = 30, hjust=1)) +
    scale_y_continuous("Temperature (ºC)", breaks=seq(0,50,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
    geom_line(aes(x = ts, y = vwc * 100,  linetype = "Volumetric Water Content (%)"), col="blue", show.legend	= TRUE) +
    scale_linetype_manual(NULL, values = 1) +
    labs(x=expression(''), colour = "Temperature (ºC)")
plot_temp_and_humidity
#ggsave('output/VWCytempAgg-selected-period-Corbalan.png', width = 15, height = 10)
```

Calculations of useful aggregation data for climate:
  * soil moisture: max VWC, range VWC, mean of VWC (misleading)
  * temp: Mean of daily temp, minimum daily temp and maximum daily temp, range temp and interquartile temp (Q3-Q1). We will use soil temperature as it's used on many other studies and it isn't so much influenced by sun irradiation and night/day differences.

```{r}
clim.daily <- db.env %>% mutate (date = date(ts)) %>% group_by(date) %>% summarise(max.temp = max(surface.temp), min.temp = min(surface.temp), mean.temp = mean(surface.temp), sd.temp = sd(surface.temp), range.vwc = max(vwc) - min(vwc), max.vwc = max(vwc), mean.vwc = mean(vwc), range.temp = max.temp - min.temp, interquartil.temp = as.numeric(quantile(surface.temp, prob=c(.75)) - quantile(surface.temp, prob=c(.25)) ) )
summary(clim.daily)
```

# Correlations

## Amplitude ~ climate

Here we show a plot with VWC vs amplitude per class

```{r}
ggplot(data = amplitude.df, mapping = aes(x = date, y = mean, col = class)) +
  ggtitle('Compare soil moisture with trees amplitude (linked to activity)') +
  geom_line() +
  scale_y_continuous("Amplitude (um)", breaks=seq(0,100,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = max.vwc * 100, linetype = "Max Volumetric Water Content (%)"), col="blue")
```

```{r}
ggplot(data = amplitude.df, mapping = aes(x = date, y = mean, col = class)) +
  ggtitle('Compare soil moisture with trees amplitude (linked to activity)') +
  geom_line() +
  scale_y_continuous("Amplitude (um)", breaks=seq(0,100,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = range.vwc * 100, linetype = "Max - Min Volumetric Water Content (%)"), col="blue")
```

Following, we will explore correlation within microclimate variables (Soil Moisture and temperature) and daily seasonal amplitude.

## Amplitude ~ Temperature
Explore correlation within temperature and daily seasonal amplitude (Not significant)

```{r}
cor.test(
  #specify the two variables to correlate
  amplitude.P_D$mean, clim.daily$interquartil.temp,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
find_Max_CCF(amplitude.P_D$mean, clim.daily$interquartil.temp)
```

## Amplitude ~ Soil Moisture
Explore cross-correlation within Soil Moisture and daily seasonal amplitude

### for all trees and classes:

```{r}
ccf (amplitude.all$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
sort_CCF_values(amplitude.all$mean, clim.daily$max.vwc)
```

### for Quercus Ilex:

Plot VWC and daily amplitude for Qi:

```{r}
plot(clim.daily$max.vwc, amplitude.Qi$mean)
```

```{r}
plot(log10(clim.daily$max.vwc), log10(amplitude.Qi$mean))
```

Test correlation:
```{r}
cor.test(
  #specify the two variables to correlate
  clim.daily$max.vwc, amplitude.Qi$mean,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
hist(log10(amplitude.Qi$mean+1))
```
```{r}
hist(log10(clim.daily$max.vwc+1))
```
Standard correlations using different methods:
```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


Cross-correlations:
```{r}
ccf (amplitude.Qi$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.Qi$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
sort_CCF_values(amplitude.Qi$mean, clim.daily$max.vwc)
sort_CCF_values(log10(amplitude.Qi$mean), log10(clim.daily$max.vwc))
```

### For P_ND: amplitude ~ VWC 

Exploring plots VWC ~ daily amplitude:

```{r}
plot(clim.daily$max.vwc, amplitude.P_ND$mean)
# plot(lm(log10(clim.daily$mean.vwc) ~ log10(amplitude.P_ND$mean)))
```


```{r}
plot(log10(clim.daily$max.vwc), log10(amplitude.P_ND$mean))
# plot(lm(clim.daily$mean.vwc ~ amplitude.P_ND$mean))
```


```{r}
join.df <- full_join(clim.daily, amplitude.P_ND, by="date")
  ggplot(data = join.df, aes(log10(max.vwc), log10(mean))) +
  # black points graph
  geom_point() +
  # add correlation with errors and blue color
  stat_smooth(method = 'lm',
              method.args = list(start= c(a = 1,b=1)),
              se=T, color = "blue") +
   # theme
  theme_classic() +
  # add labels
  labs( x = "Max Soil Moisture per day - Log",
        y = "amplitude variations (um) - Log")
```
Exploring correlations:
```{r}
cor.test(log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc),
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


```{r}
ccf (amplitude.P_ND$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
sort_CCF_values(amplitude.P_ND$mean, clim.daily$max.vwc)
sort_CCF_values(log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc))
```

## With P_D

```{r}
cor.test(log10(amplitude.P_D$mean), log10(clim.daily$max.vwc),
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_D$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_D$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


```{r}
ccf (amplitude.P_D$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.P_D$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
sort_CCF_values(amplitude.P_D$mean, clim.daily$max.vwc)
sort_CCF_values(log10(amplitude.P_D$mean), log10(clim.daily$max.vwc))
```


# Seasonality ~ climate

Now, let's see if we can see a relation within Seasonality and climate.

## All trees
Let's add up all trees seasonality to a matrix with env data and explore the correlations:
```{r}
matrix <- inner_join(seasonalities.P_D.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix
```

```{r}
library("correlation")

results <- correlation(matrix)
results
```

```{r}
library(see)

results %>%
  summary(redundant = TRUE) %>%
  plot()
```
Now using spearman method:
```{r}
results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```
Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

## By class

Let's test correlations within temp and seasonality for each different class:

### Quercus Ilex
```{r}
matrix <- inner_join(seasonalities.qi.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "pearson")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

We found that after 6 instances of 15 mins (1h and half), there is 0.70 inverse correlation within surface temperature and Quercus ilex maximum swelling.

### Non-Declining Pinus
```{r}
matrix <- inner_join(seasonalities.P_ND.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

### Declining Pinus
```{r}
matrix <- inner_join(seasonalities.P_D.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```


# Exploring wavelets

Here we test if there's another correlation with a longer frequency than the daily frequency between temperature and daily seasonality. We could think that, in warm periods (days, weeks, months,...) like the summer season we have a correlation with an specific seasonality.
I would expect to have a year correlation too (every summer we have similar tree), but unfortunately we don't have enough data to test this yet.

Next steps take a long while so,
```{r}
stop('Wavelet analysis below')
```

## For all trees: surface.temp ~ seasonality

```{r}
surface.temp = db.env %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts)) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(db.env[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, surface.temp)
           
any(is.na(surface.temp))
summary(surface.temp)

seasonality = seasonalities.agg %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts), value = mean) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(seasonalities.agg[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, value)
seasonality <- data.frame(seasonality)

any(is.na(seasonality))
summary(seasonality)

# wavelet analysis (this will take a while)
wavelet = wtc(surface.temp, seasonality,
              max.scale = 32,
              #display the progress bar
              quiet = F)

# wavelet plot
par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(wavelet,
     #legend colors
     plot.cb = TRUE,
     #lag phase (lines)
     plot.phase = TRUE, 
     ylab = "Frequency (days)",
     xlab = "Time (days)") 
```

## By class: surface.temp ~ seasonality

For Quercus:
```{r}
seasonality = seasonalities.qi.agg %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts), value = mean) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(seasonalities.agg[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, value)
seasonality <- data.frame(seasonality)

any(is.na(seasonality))
summary(seasonality)

# wavelet analysis (this will take a while)
wavelet = wtc(surface.temp, seasonality,
              max.scale = 32,
              #display the progress bar
              quiet = F)

# wavelet plot
par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(wavelet,
     #legend colors
     plot.cb = TRUE,
     #lag phase (lines)
     plot.phase = TRUE, 
     ylab = "Frequency (days)",
     xlab = "Time (days)") 
```

For Non-Declining pines:
```{r}
seasonality = seasonalities.P_ND.agg %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts), value = mean) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(seasonalities.agg[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, value)
seasonality <- data.frame(seasonality)

any(is.na(seasonality))
summary(seasonality)

# wavelet analysis (this will take a while)
wavelet = wtc(surface.temp, seasonality,
              max.scale = 32,
              #display the progress bar
              quiet = F)

# wavelet plot
par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(wavelet,
     #legend colors
     plot.cb = TRUE,
     #lag phase (lines)
     plot.phase = TRUE, 
     ylab = "Frequency (days)",
     xlab = "Time (days)") 
```

For Declining pines:
```{r}
seasonality = seasonalities.P_D.agg %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts), value = mean) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(seasonalities.agg[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, value)
seasonality <- data.frame(seasonality)

any(is.na(seasonality))
summary(seasonality)

# wavelet analysis (this will take a while)
wavelet = wtc(surface.temp, seasonality,
              max.scale = 32,
              #display the progress bar
              quiet = F)

# wavelet plot
par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(wavelet,
     #legend colors
     plot.cb = TRUE,
     #lag phase (lines)
     plot.phase = TRUE, 
     ylab = "Frequency (days)",
     xlab = "Time (days)") 
```

