---
title: "Seasonality-analysis"
author: "Abel Serrano Juste"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages:
```{r}
source('lib-dendro.R')
library(glue)
library(tidyverse)
library (ggplot2)
library(lubridate)

library(imputeTS)
library(anomalize) # Decomposition
library(stats) # Decomposition
library(ggpubr) # ggarange
library(zoo)
```

Set global vars
```{r}
### DEFINE GLOBAL VARS ###
PATH = dirname(rstudioapi::getSourceEditorContext()$path)
print(PATH)
setwd(PATH)

PLACE = 'Miedes'
DATA_DIR = 'processed/Miedes-last'
ENV_DIR = 'processed/Miedes-env-processed'
```

Select dendros for this analysis
```{r}
selected_dendros <- c("92222156", "92222169", "92222175", "92222157", "92222154", "92222170", "92222173", "92222180", "92222155", "92222163", "92222171", "92222161", "92222164")
```


Load dataset:
```{r}
list_files <- list.files(file.path(".",DATA_DIR), pattern="*.csv$", full.names=TRUE)
db<-read.all.processed(list_files)
```

## CLEAN & PREPARE DATA ###

keep data of selected dendros only
```{r}
db = db %>% filter(db$series %in% selected_dendros)
str(db)
```

select from May 2023 to Sept 2023
```{r}
# Set initial and final date for analysis
ts_start <- "2023-05-01 09:00:00" # 2 days after installation
ts_end <-"2023-09-10 07:00:00" # last timestamp of downloaded data
db <- reset.initial.values(db, ts_start, ts_end)
```

# INSPECT DATA
```{r}
str(db)
head(db)
tail(db)
```

# Data imputation

Missing data
```{r}
statsNA(db$value)
```

Let's fill it through interpolation
```{r}
db$value <- db$value %>% 
            na_interpolation(option = "spline")
```

# Time Series decomposition

## Define useful functions

Define plot_decompose function which uses basic time series decomposition
```{r}
## Function decompose
plot_decompose <- function (db, name) {
  # decompose a time series (has to be in tibble format)
  decompose = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select varaible to decompose
    value)
  
  # plot time series
  p_ts = ggplot (decompose, aes(x = ts, y = observed)) +
    ggtitle(glue("Time series decomposition for {name}")) +
    geom_line (col = "black") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = NULL,
             y = "Original data (um)")
  
  # plot trend
  p_trend = ggplot (decompose, aes(x = ts, y = trend)) +
    geom_line (col = "#D55E00") + 
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    geom_hline(yintercept = 0, linetype='dotted', col = 'red') +
    theme_classic() +
    labs ( x = NULL,
             y = "Trend")
  
  # plot season
  p_season = ggplot (decompose, aes(x = ts, y = season)) +
    geom_line (col = "#E69F00") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = NULL,
             y = "Season")
  
  # plot remainder
  p_remainder = ggplot (decompose, aes(x = ts, y = remainder)) +
    geom_line (col = "#F0E442") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = "Period (month)",
             y = "Remainder")
  
  # plot all together
  ggarrange (p_ts, p_trend, p_season, p_remainder, ncol = 1)
}

# Save the plot of decompose_ts
save_plot_decompose <- function (dat, name) {
  plot_decompose(dat, name) %>% ggexport(filename = glue('output/{name}-decomp-anomalize.png'),  width = 4500, height = 3000, res = 300)
}
```

Define amplitude function
```{r}
calculate_amplitude <- function(dat) {
  dat %>% mutate(date = date(ts)) %>% group_by(date) %>% summarize(max = max(value), min = min(value)) %>% mutate(ampl = max-min)
}

save_plot_amplitude <- function(dat.ampl, name) {
  ggplot(data = dat.ampl, mapping = aes(x = date, y = ampl)) + geom_line()
  ggsave(glue('output/{name}-amplitude.png'), width = 15, height = 10)
}
```

Define stl function
```{r}
plot_stl <- function (stl.out) {
  summary(stl.out)
  plot(stl.out)
}

save_plot_stl <- function (stl.out, name) {
  png(glue('output/{name}-stl.png'), width=15, height=10, units="in", res=300)
  plot_stl(stl.out)
  dev.off()
}

# Now plotting only seasonality but with dates in the x-axis
plot_seasonality <- function(stl.out, name) {
  
  seasonality <- stl.out$time.series[,1]
  timestamps <- seq(from = as.POSIXct(ts_start, tz='Madrid/Spain'), by = "15 min", length.out = length(dat.ts))
  zoo_data <- zoo(seasonality, order.by = timestamps)
  plot(zoo_data, xaxt = "n", type = "l", xlab = "", ylab = "Value", main = glue("Time Series by Month-Year for {name}"))
  axis(1, at = time(zoo_data), labels = format(time(zoo_data), "%Y-%m"))
  
  # Add x-axis label
  mtext("Month-Year", side = 1, line = 3)
}

save_plot_seasonality_stl <- function(stl.out, name) {
  png(glue('output/{name}-seasonality-stl.png'), width=15, height=10, units="in", res=300)
  plot_seasonality(stl.out, name)
  dev.off()
}
```


# Calculate mean Seasonality
Mean seasonality for all trees:
```{r}
seasonalities <- data.frame()
dendro.series <- unique(db$series)

for (dendro.no in dendro.series) {
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  dat.ts <- ts(data = dat$value, frequency = 96)
  
  stl.out = stl(dat.ts, s.window = 25, t.window = 673)

  seasonality <- stl.out$time.series[,1]
  
  aux <- data.frame(
    value = as.numeric(seasonality),
    series = as.factor(dendro.no),
    ts = dat$ts
  )

  seasonalities <- rbind.data.frame(seasonalities, aux)
}
summary(seasonalities)

seasonalities.agg <- seasonalities %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.agg)
```

```{r}
ggplot( data = seasonalities.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for all trees")
ggsave('output/mean-seasonalities.png', width = 15, height = 10)
```


# Calculate Seasonality by class

Calculate mean of the seasonality for Qi, PD and P_ND
```{r}
Qi = TreeList %>% filter(class == "Quercus", series %in% unique(db$series)) %>% pull(series)
P_D = TreeList %>% filter(class == "D", series %in% unique(db$series)) %>% pull(series)
P_ND = TreeList %>% filter(class == "ND", series %in% unique(db$series)) %>% pull(series)
```

Mean seasonality for Quercus Ilex:
```{r}
seasonalities.qi <- data.frame()

for (dendro.no in Qi) {
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  dat.ts <- ts(data = dat$value, frequency = 96)
  
  stl.out = stl(dat.ts, s.window = 25, t.window = 673)

  seasonality <- stl.out$time.series[,1]
  
  aux <- data.frame(
    value = as.numeric(seasonality),
    series = as.factor(dendro.no),
    ts = dat$ts
  )

  seasonalities.qi <- rbind.data.frame(seasonalities.qi, aux)
}
summary(seasonalities.qi)

seasonalities.qi.agg <- seasonalities.qi %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.qi.agg)
```

```{r}
ggplot( data = seasonalities.qi.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Quercus Ilex")
ggsave('output/mean-seasonality-Qi.png', width = 15, height = 10)
```

Mean seasonality for Non-Declining Pinus
```{r}
seasonalities.P_ND <- data.frame()

for (dendro.no in P_ND) {
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  dat.ts <- ts(data = dat$value, frequency = 96)

  stl.out = stl(dat.ts, s.window = 25, t.window = 673)

  seasonality <- stl.out$time.series[,1]
  
  aux <- data.frame(
    value = as.numeric(seasonality),
    series = as.factor(dendro.no),
    ts = dat$ts
  )

  seasonalities.P_ND <- rbind.data.frame(seasonalities.P_ND, aux)
}
summary(seasonalities.P_ND)

seasonalities.P_ND.agg <- seasonalities.P_ND %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.P_ND.agg)
```

```{r}
ggplot( data = seasonalities.P_ND.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Non-Declining Pinus")
ggsave('output/mean-seasonality-P_ND.png', width = 15, height = 10)
```

Mean seasonality for Declining Pinus
```{r}
seasonalities.P_D <- data.frame()

for (dendro.no in P_D) {
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  dat.ts <- ts(data = dat$value, frequency = 96)
  
  stl.out = stl(dat.ts, s.window = 25, t.window = 673)

  seasonality <- stl.out$time.series[,1]
  
  aux <- data.frame(
    value = as.numeric(seasonality),
    series = as.factor(dendro.no),
    ts = dat$ts
  )

  seasonalities.P_D <- rbind.data.frame(seasonalities.P_D, aux)
  remove(aux)
}
summary(seasonalities.P_D)

seasonalities.P_D.agg <- seasonalities.P_D %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.P_D.agg)
```

```{r}
ggplot( data = seasonalities.P_D.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Declining Pinus")
ggsave('output/mean-seasonality-P_D.png', width = 15, height = 10)
```

# Calculate amplitude for all trees

```{r}
TreeList <- read.table("TreeList.txt", header=T)

# First, define dendro number
column_names <- c("date", "min", "max", "ampl", "class")
amplitude.df = data.frame(matrix(nrow = 0, ncol = length(column_names)))
colnames(amplitude.df) <- column_names
for (dendro.no in selected_dendros) {
  
  class = (TreeList[TreeList$series == dendro.no,])$class
  name = paste(dendro.no, class, sep="-")
  
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  
  # Basic ts decomposition
  # save_plot_decompose(dat, name)
  
  # Max - min daily amplitude
  dat.ampl <- calculate_amplitude(dat)
  aux <- cbind(ampl = dat.ampl, class=class) %>% rename (date = ampl.date, min = ampl.min, max = ampl.max, ampl = ampl.ampl)
  amplitude.df = rbind.data.frame(amplitude.df, aux)
}
```

With all daily amplitudes, calculate mean of each one by class

```{r}
amplitude.df = amplitude.df %>% group_by(date, class) %>% summarise(mean = mean(ampl))
amplitude.df
```

Create one different dataframe for every class
```{r}
amplitude.Qi <- amplitude.df[amplitude.df$class == "Quercus",]
amplitude.P_ND <- amplitude.df[amplitude.df$class == "ND",]
amplitude.P_D <- amplitude.df[amplitude.df$class == "D",]
```


# Climate data

Importing environmental data and keeping the one sensor with all valid data
```{r}
db.env <- read.env.proc(file.path(PATH,ENV_DIR,'proc-env.csv'))
db.env <- db.env[db.env$series == "94231940",]
```

```{r}
plot_ly(db.env, x = ~ts, y = ~vwc, color = ~series, type = 'scatter', mode = 'lines')
```


Filter env data to period of study
```{r}
db.env <- db.env[which(db.env$ts>=ts_start & db.env$ts<=ts_end),]
summary(db.env)
```

Plot env data:
```{r}
plot_temp_and_humidity <-
  ggplot(data = db.env, aes(x=ts, y=soil.temp)) +
    ggtitle("Temperature and humidity for Miedes (2022-2023)") +
    geom_line(aes(colour = soil.temp)) +
    theme_bw() +
    scale_colour_gradient(low = "light blue", high = "red") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme(axis.text.x = element_text(angle = 30, hjust=1)) +
    scale_y_continuous("Temperature (ºC)", breaks=seq(0,50,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
    geom_line(aes(x = ts, y = vwc * 100,  linetype = "Volumetric Water Content (%)"), col="blue", show.legend	= TRUE) +
    scale_linetype_manual(NULL, values = 1) +
    labs(x=expression(''), colour = "Temperature (ºC)")
plot_temp_and_humidity
```

Calculations of useful aggregation data for climate:
  * soil moisture: max VWC, range VWC, mean of VWC (misleading)
  * temp: Mean of daily temp, minimum daily temp and maximum daily temp, range temp and interquartile temp (Q3-Q1). We will use soil temperature as it's used on many other studies and it isn't so much influenced by sun irradiation and night/day differences.

```{r}
clim.daily <- db.env %>% mutate (date = date(ts)) %>% group_by(date) %>% summarise(max.temp = max(soil.temp), min.temp = min(soil.temp), mean.temp = mean(soil.temp), sd.temp = sd(soil.temp), range.vwc = max(vwc) - min(vwc), max.vwc = max(vwc), mean.vwc = mean(vwc), range.temp = max.temp - min.temp, interquartil.temp = as.numeric(quantile(soil.temp, prob=c(.75)) - quantile(soil.temp, prob=c(.25)) ) )
summary(clim.daily)
```


# Correlations

## Amplitude ~ climate

Here we show a plot with VWC vs amplitude per class

```{r}
ggplot(data = amplitude.df, mapping = aes(x = date, y = mean, col = class)) +
  ggtitle('Compare soil moisture with trees amplitude (linked to activity)') +
  geom_line() +
  scale_y_continuous("Amplitude (um)", breaks=seq(0,100,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = max.vwc * 100, linetype = "Max Volumetric Water Content (%)"), col="blue")
```

```{r}
ggplot(data = amplitude.df, mapping = aes(x = date, y = mean, col = class)) +
  ggtitle('Compare soil moisture with trees amplitude (linked to activity)') +
  geom_line() +
  scale_y_continuous("Amplitude (um)", breaks=seq(0,100,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = range.vwc * 100, linetype = "Max - Min Volumetric Water Content (%)"), col="blue")
```


Following, we will Explore correlation within microclimate variables (Soil Moisture and temperature) and daily seasonal amplitude.

## for Quercus Ilex:

Plot VWC and daily amplitude for Qi:

```{r}
plot(clim.daily$mean.vwc, amplitude.Qi$mean)
```


```{r}
plot(log10(clim.daily$mean.vwc), log10(amplitude.Qi$mean))
```

Test correlation:
```{r}
cor.test(
  #specify the two variables to correlate
  clim.daily$max.vwc, amplitude.Qi$mean,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
hist(log10(amplitude.Qi$mean+1))
```
```{r}
hist(log10(clim.daily$mean.vwc+1))
```

Explore correlation within temperature and daily seasonal amplitude (Not significant)
```{r}
cor.test(
  #specify the two variables to correlate
  amplitude.P_D$mean, clim.daily$interquartil.temp,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
find_Max_CCF(amplitude.P_D$mean, clim.daily$interquartil.temp)
```

Explore cross-correlation within Soil Moisture and daily seasonal amplitude

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```



```{r}
ccf (amplitude.Qi$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
find_Max_CCF = function(x,y) {
 # run cross-correlation function
 ccf = ccf(x, y, plot = FALSE, na.action = na.pass) 
 # build a dataset with lag times and correlation coefficients
 res_cc = data.frame(lag = ccf$lag[,,1], cor = ccf$acf[,,1]) 
 max = res_cc[which.max(abs(res_cc$cor)),] 
 # return only the data of interest
 return(max) 
} 

find_Max_CCF(amplitude.Qi$mean, clim.daily$max.vwc)
```

## With P_ND

Plot VWC and daily amplitude:

```{r}
join.df <- full_join(clim.daily, amplitude.P_ND, by="date")
  ggplot(data = join.df, aes(log10(max.vwc), log10(mean))) +
  # black points graph
  geom_point() +     
  # add correlation with errors and blue color
  stat_smooth(method = 'lm', 
              method.args = list(start= c(a = 1,b=1)),
              se=T, color = "blue") +
   # theme
  theme_classic() +
  # add labels
  labs( x = "Air Light (lux)",
        y = expression("Range in"~O[2]~(mg~L^-1)))
```


```{r}
plot(clim.daily$max.vwc, amplitude.P_ND$mean)
# plot(lm(log10(clim.daily$mean.vwc) ~ log10(amplitude.P_ND$mean)))
```


```{r}
plot(log10(clim.daily$max.vwc), log10(amplitude.P_ND$mean))
# plot(lm(clim.daily$mean.vwc ~ amplitude.P_ND$mean))
```

```{r}
cor.test(log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc),
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


```{r}
ccf (log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
find_Max_CCF(log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc))
```

## With P_D

```{r}
cor.test(log10(amplitude.P_D$mean), log10(clim.daily$max.vwc),
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_D$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_D$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


```{r}
ccf (log10(amplitude.P_D$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
find_Max_CCF(log10(amplitude.P_D$mean), log10(clim.daily$max.vwc))
```


# Seasonality ~ climate

Now, let's see if we can see a relation within Seasonality and climate.

## All trees
Let's add up all trees seasonality to a matrix with env data and explore the correlations:
```{r}
matrix <- inner_join(seasonalities.P_D.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix
```

```{r}
library("correlation")

results <- correlation(matrix)
results
```

```{r}
library(see)

results %>%
  summary(redundant = TRUE) %>%
  plot()
```
Now using spearman method:
```{r}
results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```
Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

## By class

Let's test correlations within temp and seasonality for each different class:

### Quercus Ilex
```{r}
matrix <- inner_join(seasonalities.qi.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "pearson")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

We found that after 6 instances of 15 mins (1h and half), there is 0.70 inverse correlation within surface temperature and Quercus ilex maximum swelling.

### Non-Declining Pinus
```{r}
matrix <- inner_join(seasonalities.P_ND.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

### Declining Pinus
```{r}
matrix <- inner_join(seasonalities.P_D.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```


## Explore wavelets

### All trees

```{r}

```

### By class
