---
title: "Seasonality-analysis"
author: "Abel Serrano Juste"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages:
```{r}
source('lib-dendro.R')
source('lib-ts-analysis.R')
source('correlations.R')

library(glue)
library(tidyverse)
library(ggplot2)
theme_set(theme_bw())
library(lubridate)
library(hms)

library(imputeTS)
# library(biwavelet)
```

Set global vars
```{r}
### DEFINE GLOBAL VARS ###
PATH = dirname(rstudioapi::getSourceEditorContext()$path)
print(PATH)
setwd(PATH)

PLACE = 'Miedes'
DATA_DIR = 'processed/Miedes-last'
ENV_DIR = 'processed/Miedes-env-processed'

TreeList <- read.table("TreeList.txt", header=T)

ALL_PERIOD_ST <- "May to October"
```

Select dendros for this analysis (all if FALSE)
```{r}
selected_dendros <- c("92222154", "92222155", "92222156", "92222157", "92222161", "92222163", "92222164", "92222169", "92222170", "92222171", "92222173", "92222175", "92222180")
# selected_dendros <- c()
```

Select TMS serial no for this analysis
```{r}
SELECTED_TMS <- "94231935"
```

Load dataset:
```{r}
list_files <- list.files(file.path(".",DATA_DIR), pattern="*.csv$", full.names=TRUE)
db<-read.all.processed(list_files)
```

## CLEAN & PREPARE DATA ###

keep data of selected dendros only
```{r}
if (length(selected_dendros) > 0) {db = db %>% filter(db$series %in% selected_dendros)}
str(db)
```

select from May 2023 to November 2023
```{r}
# Set initial and final date for analysis
ts_start <- "2023-05-01 09:00:00" # from first of May
ts_end <-"2023-11-01 07:00:00" # to 1st of november
db <- reset.initial.values(db, ts_start, ts_end)
```

Add group data
```{r}
db <- db %>% mutate (
  class = case_when(
    series %in% Qi ~ factor("Quercus"),
    series %in% P_D ~ factor("D"),
    series %in% P_ND ~ factor("ND"),
    .default = NA
  )
)
```


# INSPECT DATA
```{r}
str(db)
head(db)
tail(db)
```

# Data imputation

Missing data
```{r}
statsNA(db$value)
```

Let's fill it through interpolation
```{r}
db$value <- db$value %>% 
            na_interpolation(option = "spline")
```

Set ggplot label options, themes & vars:
```{r}
theme_set( theme_bw() +
          theme(
            axis.text=element_text(size=12),
            legend.text=element_text(size=10),
            axis.title=element_text(size=14),
            plot.title = element_text(hjust = 0.5, face = "bold")
          ))

every_hour_labels = format(lubridate::parse_date_time(hms(hours = 0:23), c('HMS', 'HM')), '%H:%M')
labels = c("Quercus", "Declining Pines", "Non-Declining Pines")

n.per.class <- db %>% group_by(class) %>% summarise (n = n_distinct(series)) %>% pivot_wider(names_from = class, values_from = n)

c_labels = c(glue("{labels[1]} ({n.per.class$Quercus})"), glue("{labels[2]} ({n.per.class$D})"), glue("{labels[3]} ({n.per.class$ND})"))
c_values = c(Quercus = "purple", D = "darkred", ND = "darkorange")
```

# Time Series decomposition

# Calculate mean Seasonality
Mean seasonality for all trees:
```{r}
seasonalities <- calculate_stl_seasonalities(db, unique(db$series))
summary(seasonalities)

seasonalities.agg <- seasonalities %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.agg)
```

```{r}
ggplot( data = seasonalities.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for all trees")
# # ggsave(glue('output/{PLACE}-mean-seasonalities-all.png'), width = 15, height = 10)
```

# Calculate Seasonality by class

Calculate mean of the seasonality for Qi, PD and P_ND
```{r}
Qi = TreeList %>% filter(class == "Quercus", series %in% unique(db$series)) %>% pull(series)
P_D = TreeList %>% filter(class == "D", series %in% unique(db$series)) %>% pull(series)
P_ND = TreeList %>% filter(class == "ND", series %in% unique(db$series)) %>% pull(series)
```

```{r}
every_hour_labels = format(lubridate::parse_date_time(hms(hours = 0:23), c('HMS', 'HM')), '%H:%M')
```

## Quercus
Mean seasonality for Quercus Ilex:
```{r}
seasonalities.qi <- calculate_stl_seasonalities(db, Qi)
summary(seasonalities.qi)

seasonalities.qi.agg <- seasonalities.qi %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.qi.agg)
```

```{r}
ggplot( data = seasonalities.qi, mapping = aes(x=ts, y = value, color = series)) + 
  geom_line(alpha = 0.5) +
  ggtitle(glue('Quercus Ilex seasonalities by dendrometer - {PLACE} '))
# # ggsave(glue('output/{PLACE}-seasonality-Qi.png'), width = 15, height = 10)
```

```{r}
ggplot( data = seasonalities.qi.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Quercus Ilex")
# # ggsave(glue('output/{PLACE}-mean-seasonality-Qi.png'), width = 15, height = 10)
```

### Aggregating in one day, by removing time-series dimension

removing the time-series dimension, mean of one day for all period:
```{r}
seasons.qi.allperiod = seasonalities.qi.agg %>% mutate(timeOfDay = as_hms(ts)) %>% summarise(meanSeasonalityTime = mean(mean), SE_SeasonalityTime = sd(sd)/sqrt(n()), .by = timeOfDay)
seasons.qi.allperiod
plot_day_seasonality(seasons.qi.allperiod, "Quercus Ilex", PLACE, ALL_PERIOD_ST)
# # ggsave(glue('output/{PLACE}_aggoneday-allperiod-seasonalities-Qi.png'), width = 15, height = 10)
```

removing the time-series dimension, mean of one day for June-July:
```{r}
seasons.qi.jun_jul = seasonalities.qi.agg %>% filter( (ts >= as.Date("2023-06-01") ) & (ts < as.Date("2023-08-01")) ) %>%  mutate(timeOfDay = as_hms(ts)) %>% summarise(meanSeasonalityTime = mean(mean), SE_SeasonalityTime = sd(sd)/sqrt(n()), .by = timeOfDay)
seasons.qi.jun_jul

plot_day_seasonality(seasons.qi.jun_jul, "Quercus Ilex", PLACE, "June to July 2023")
# # ggsave(glue('output/{PLACE}_aggoneday-june-july-seasonalities-Qi.png'), width = 15, height = 10)
```


Now, the two plots above together in one plot:

```{r}
seasons.qi.jun_jul$period <- "June-July"
seasons.qi.allperiod$period <- "All"
seasons.periods.joined <- rbind.data.frame(seasons.qi.allperiod, seasons.qi.jun_jul)
```


```{r}
plot_day_seasonalities_periods(seasons.periods.joined, "Quercus Ilex", PLACE)
# # ggsave(glue('output/{PLACE}_aggoneday-bothperiods-seasonalities-Qi.png'), width = 15, height = 10)
```


```{r}
# Another way to do it but using two different df sources
# ggplot() +
#   ggtitle("Aggregated mean in one day for all period And for period june-july") +
#   geom_line(data = seasons.qi.allperiod, mapping = aes(x=timeOfDay, y = meanSeasonalityTime), col = "blue") + 
#   geom_line(data = seasons.qi.allperiod, aes (x = timeOfDay, y = (meanSeasonalityTime + SE_SeasonalityTime)), col = "red", alpha = 0.5, linetype = "dashed", show.legend = F) +
#   geom_line(data = seasons.qi.allperiod, aes (x = timeOfDay, y = (meanSeasonalityTime - SE_SeasonalityTime)), col = "red", alpha = 0.5, linetype = "dashed", show.legend = F) +
#   geom_line(data = seasons.qi.jun_jul, mapping = aes(x = timeOfDay, y = meanSeasonalityTime), col= "purple") +
#   geom_line(data = seasons.qi.jun_jul, aes (x = timeOfDay, y = (meanSeasonalityTime + SE_SeasonalityTime)), col = "red", alpha = 0.5, linetype = "dashed", show.legend = F) +
#   geom_line(data = seasons.qi.jun_jul, aes (x = timeOfDay, y = (meanSeasonalityTime - SE_SeasonalityTime)), col = "red", alpha = 0.5, linetype = "dashed", show.legend = F) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   scale_x_time(breaks = seq(0, 85500, by = 3600), labels = every_hour_labels) +
#   labs(x = "Hour of the day", y = "Micrometers of Daily seasonality (um)")
```


## Non-Declining Pines
Mean seasonality for Non-Declining Pines
```{r}
seasonalities.P_ND <- calculate_stl_seasonalities(db, P_ND)
summary(seasonalities.P_ND)

seasonalities.P_ND.agg <- seasonalities.P_ND %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.P_ND.agg)
```

```{r}
ggplot( data = seasonalities.P_ND, mapping = aes(x=ts, y = value, color = series)) + 
  geom_line(alpha = 0.5) +
  ggtitle(glue('Non-Declining Pinus seasonalities by dendrometer - {PLACE} '))
# # ggsave(glue('output/{PLACE}-seasonality-P_ND.png'), width = 15, height = 10)
```

```{r}
ggplot( data = seasonalities.P_ND.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Non-Declining Pinus")
# # ggsave(glue('output/{PLACE}-mean-seasonality-P_ND.png'), width = 15, height = 10)
```


### Non-Declining Pines: Aggregating in one day, by removing time-series dimension

removing the time-series dimension, mean of one day for all period:
```{r}
seasons.P_ND.allperiod = seasonalities.P_ND.agg %>% mutate(timeOfDay = as_hms(ts)) %>% summarise(meanSeasonalityTime = mean(mean), SE_SeasonalityTime = sd(sd)/sqrt(n()), .by = timeOfDay)
seasons.P_ND.allperiod
plot_day_seasonality(seasons.P_ND.allperiod, "Non-Declining Pines", PLACE, ALL_PERIOD_ST)
# # ggsave(glue('output/{PLACE}_aggoneday-allperiod-seasonalities-P_ND.png'), width = 15, height = 10)
```

removing the time-series dimension, mean of one day for June-July:
```{r}
seasons.P_ND.jun_jul = seasonalities.P_ND.agg %>% filter( (ts >= as.Date("2023-06-01") ) & (ts < as.Date("2023-08-01")) ) %>%  mutate(timeOfDay = as_hms(ts)) %>% summarise(meanSeasonalityTime = mean(mean), SE_SeasonalityTime = sd(sd)/sqrt(n()), .by = timeOfDay)
seasons.P_ND.jun_jul

plot_day_seasonality(seasons.P_ND.jun_jul, "Non-Declining Pines", PLACE, "June to July 2023")
# # ggsave(glue('output/{PLACE}_aggoneday-june-july-seasonalities-P_ND.png'), width = 15, height = 10)
```


Now, the two plots above together in one plot:

```{r}
seasons.P_ND.jun_jul$period <- "June-July"
seasons.P_ND.allperiod$period <- "All"
seasons.periods.joined <- rbind.data.frame(seasons.P_ND.allperiod, seasons.P_ND.jun_jul)
```


```{r}
plot_day_seasonalities_periods(seasons.periods.joined, "Non-Declining Pines", PLACE)
# # ggsave(glue('output/{PLACE}_aggoneday-bothperiods-seasonalities-P_ND.png'), width = 15, height = 10)
```


## Declining Pines
Mean seasonality for Declining Pines
```{r}
seasonalities.P_D <- calculate_stl_seasonalities(db, P_D)
summary(seasonalities.P_D)

seasonalities.P_D.agg <- seasonalities.P_D %>% group_by(ts) %>% summarise(mean = mean(value), sd = sd(value))
summary(seasonalities.P_D.agg)
```

```{r}
ggplot( data = seasonalities.P_D, mapping = aes(x=ts, y = value, color = series)) + 
  geom_line(alpha = 0.5) +
  ggtitle(glue('Declining Pinus seasonalities by dendrometer - {PLACE} '))
# # ggsave(glue('output/{PLACE}-seasonality-P_D.png'), width = 15, height = 10)
```

```{r}
ggplot( data = seasonalities.P_D.agg, mapping = aes(x=ts, y = mean)) + 
  geom_line() + 
  ggtitle("Mean of seasonalities for Declining Pinus")
# # ggsave(glue('output/{PLACE}-mean-seasonality-P_D.png'), width = 15, height = 10)
```


### Declining Pines: Aggregating in one day, by removing time-series dimension

removing the time-series dimension, mean of one day for all period:
```{r}
seasons.P_D.allperiod = seasonalities.P_D.agg %>% mutate(timeOfDay = as_hms(ts)) %>% summarise(meanSeasonalityTime = mean(mean), SE_SeasonalityTime = sd(sd)/sqrt(n()), .by = timeOfDay)
seasons.P_D.allperiod
plot_day_seasonality(seasons.P_D.allperiod, "Declining Pines", PLACE, ALL_PERIOD_ST)
# # ggsave(glue('output/{PLACE}_aggoneday-allperiod-seasonalities-P_D.png'), width = 15, height = 10)
```

removing the time-series dimension, mean of one day for June-July:
```{r}
seasons.P_D.jun_jul = seasonalities.P_D.agg %>% filter( (ts >= as.Date("2023-06-01") ) & (ts < as.Date("2023-08-01")) ) %>%  mutate(timeOfDay = as_hms(ts)) %>% summarise(meanSeasonalityTime = mean(mean), SE_SeasonalityTime = sd(sd)/sqrt(n()), .by = timeOfDay)
seasons.P_D.jun_jul

plot_day_seasonality(seasons.P_D.jun_jul, "Declining Pines", PLACE, "June to July 2023")
# # ggsave(glue('output/{PLACE}_aggoneday-june-july-seasonalities-P_D.png'), width = 15, height = 10)
```

Now, the two plots above together in one plot:

```{r}
seasons.P_D.jun_jul$period <- "June-July"
seasons.P_D.allperiod$period <- "All"
seasons.periods.joined <- rbind.data.frame(seasons.P_D.allperiod, seasons.P_D.jun_jul)
```


```{r}
plot_day_seasonalities_periods(seasons.periods.joined, "Declining Pines", PLACE)
# # ggsave(glue('output/{PLACE}_aggoneday-bothperiods-seasonalities-P_D.png'), width = 15, height = 10)
```


### Plot aggregated data of each class in one plot

All study period:
```{r}
seasons.allclases.allperiod <- data.frame()
seasons.P_D.allperiod$class = "D"
seasons.P_ND.allperiod$class = "ND"
seasons.qi.allperiod$class = "Quercus"
seasons.allclases.allperiod <- rbind.data.frame(seasons.allclases.allperiod, seasons.P_D.allperiod, seasons.P_ND.allperiod, seasons.qi.allperiod)
plot_day_seasonality_byclass(seasons.allclases.allperiod, "All trees", PLACE, ALL_PERIOD_ST)
# ggsave(glue('output/{PLACE}_aggoneday-allperiod-seasonalities-alltrees-byclass.png'), width = 15, height = 10)
```

June to July:
```{r}
seasons.allclases.jun_jul <- data.frame()
seasons.P_D.jun_jul$class = "D"
seasons.P_ND.jun_jul$class = "ND"
seasons.qi.jun_jul$class = "Quercus"
seasons.allclases.jun_jul <- rbind.data.frame(seasons.allclases.jun_jul, seasons.P_D.jun_jul, seasons.P_ND.jun_jul, seasons.qi.jun_jul)
plot_day_seasonality_byclass(seasons.allclases.jun_jul, "All trees", PLACE, "Summer")
# ggsave(glue('output/{PLACE}_aggoneday-june-july-seasonalities-alltrees-byclass.png'), width = 15, height = 10)
```


# Calculate amplitude for all trees by class

```{r}
# First, let's create an amplitude.df to store the output.
column_names <- c("date", "min", "max", "ampl", "class")
amplitude.df = data.frame(matrix(nrow = 0, ncol = length(column_names)))
colnames(amplitude.df) <- column_names
for (dendro.no in unique(db$series)) {
  
  # skip if we don't have the dendro registered in TreeList
  if (! (dendro.no %in% TreeList$series)) next;
  
  class = (TreeList[TreeList$series == dendro.no,])$class
  name = paste(dendro.no, class, sep="-")
  
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  
  # Max - min daily amplitude
  dat.ampl <- calculate_amplitude(dat)
  aux <- cbind(ampl = dat.ampl, name = name, class=class) %>% rename (date = ampl.date, min = ampl.min, max = ampl.max, ampl = ampl.ampl)
  amplitude.df = rbind.data.frame(amplitude.df, aux)
}
```

With all daily amplitudes, calculate mean of each one by class

```{r}
amplitude.all = amplitude.df %>% group_by(date) %>% summarise(mean = mean(ampl), .groups = "drop")
amplitude.all
amplitude.df = amplitude.df %>% group_by(date, class) %>% summarise(mean = mean(ampl))
amplitude.df
```

Create one different dataframe for every class
```{r}
amplitude.Qi <- amplitude.df[amplitude.df$class == "Quercus",]
amplitude.P_ND <- amplitude.df[amplitude.df$class == "ND",]
amplitude.P_D <- amplitude.df[amplitude.df$class == "D",]
```

# Climate data

Importing environmental data and keeping the one sensor with all valid data
```{r}
db.env <- read.env.proc(file.path(PATH,ENV_DIR,'proc-env.csv'))
db.env <- db.env[db.env$series == SELECTED_TMS,]
```

Calculate mean among all climate sensors with all its data *complete* (no gaps)
```{r}
# statsNA(db.env$vwc)
# db.env$vwc <- db.env$vwc %>% 
#             na_interpolation(option = "spline")
```


```{r}
#library(plotly)
#plot_ly(db.env, x = ~ts, y = ~vwc, color = ~series, type = 'scatter', mode = 'lines')
```


Filter env data to period of study
```{r}
db.env <- db.env[which(db.env$ts>=ts_start & db.env$ts<=ts_end),]
summary(db.env)
```

Plot env data:
```{r}
plot_temp_and_humidity <-
  ggplot(data = db.env, aes(x=ts, y=surface.temp)) +
    ggtitle(glue("Temperature and humidity for {PLACE} (May to September 2023)")) +
    geom_line(aes(colour = surface.temp)) +
    theme_bw() +
    scale_colour_gradient(low = "light blue", high = "red") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme(axis.text.x = element_text(angle = 30, hjust=1)) +
    scale_y_continuous("Temperature (ºC)", breaks=seq(0,50,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
    geom_line(aes(x = ts, y = vwc * 100,  linetype = "Volumetric Water Content (%)"), col="blue", show.legend	= TRUE) +
    scale_linetype_manual(NULL, values = 1) +
    labs(x=expression(''), colour = "Temperature (ºC)")
plot_temp_and_humidity
# # ggsave(glue('output/{PLACE}-VWCytempAgg-selected-period.png'), width = 15, height = 10)
```

Calculations of useful aggregation data for climate:
  * soil moisture: max VWC, range VWC, mean of VWC (misleading)
  * temp: Mean of daily temp, minimum daily temp and maximum daily temp, range temp and interquartile temp (Q3-Q1). We will use soil temperature as it's used on many other studies and it isn't so much influenced by sun irradiation and night/day differences.

```{r}
clim.daily <- db.env %>% mutate (date = date(ts)) %>% group_by(date) %>% summarise(max.temp = max(surface.temp), min.temp = min(surface.temp), mean.temp = mean(surface.temp), sd.temp = sd(surface.temp), range.vwc = max(vwc) - min(vwc), max.vwc = max(vwc), mean.vwc = mean(vwc), range.temp = max.temp - min.temp, interquartil.temp = as.numeric(quantile(surface.temp, prob=c(.75)) - quantile(surface.temp, prob=c(.25)) ) )
summary(clim.daily)
```

# Correlations

## Amplitude ~ climate

Here we show a plot with VWC vs amplitude per class

```{r}
ggplot(data = amplitude.df, mapping = aes(x = date, y = mean, col = class)) +
  ggtitle(glue('Soil moisture vs daily stem amplitude difference for {PLACE}')) +
  geom_line() +
  scale_y_continuous("Amplitude (um)", breaks=seq(0,100,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = range.vwc * 100, linetype = "Max - Min Volumetric Water Content (%)"), col="blue")
```

Following, we will explore correlation within microclimate variables (Soil Moisture and temperature) and daily seasonal amplitude.

## Amplitude ~ Temperature
Explore correlation within temperature and daily stem amplitude difference (Not significant)

```{r}
cor.test(
  #specify the two variables to correlate
  amplitude.P_D$mean, clim.daily$interquartil.temp,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
find_Max_CCF(amplitude.P_D$mean, clim.daily$interquartil.temp)
```

## Amplitude ~ Soil Moisture
Explore cross-correlation within Soil Moisture and daily seasonal amplitude

### for all trees and classes:

```{r}
ccf (amplitude.all$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
sort_CCF_values(amplitude.all$mean, clim.daily$max.vwc)
```

### for Quercus Ilex:

Plot VWC and daily amplitude for Qi:

```{r}
plot(clim.daily$max.vwc, amplitude.Qi$mean)
```

```{r}
plot(log10(clim.daily$max.vwc), log10(amplitude.Qi$mean))
```

Test correlation:
```{r}
cor.test(
  #specify the two variables to correlate
  clim.daily$max.vwc, amplitude.Qi$mean,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
hist(log10(amplitude.Qi$mean+1))
```
```{r}
hist(log10(clim.daily$max.vwc+1))
```

Standard correlations using different methods:
```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.Qi$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


Cross-correlations:
```{r}
ccf (amplitude.Qi$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.Qi$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
find_Max_CCF(amplitude.Qi$mean, clim.daily$max.vwc)
sort_CCF_values(amplitude.Qi$mean, clim.daily$max.vwc)
```

### For P_ND: amplitude ~ VWC 

Exploring plots VWC ~ daily amplitude:

```{r}
plot(clim.daily$max.vwc, amplitude.P_ND$mean)
# plot(lm(log10(clim.daily$mean.vwc) ~ log10(amplitude.P_ND$mean)))
```


```{r}
plot(log10(clim.daily$max.vwc), log10(amplitude.P_ND$mean))
# plot(lm(clim.daily$mean.vwc ~ amplitude.P_ND$mean))
```


```{r}
join.df <- full_join(clim.daily, amplitude.P_ND, by="date")
  ggplot(data = join.df, aes(log10(max.vwc), log10(mean))) +
  # black points graph
  geom_point() +
  # add correlation with errors and blue color
  stat_smooth(method = 'lm',
              method.args = list(start= c(a = 1,b=1)),
              se=T, color = "blue") +
   # theme
  theme_classic() +
  # add labels
  labs( x = "Max Soil Moisture per day - Log",
        y = "amplitude variations (um) - Log")
```
Exploring correlations:
```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_ND$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


```{r}
ccf (amplitude.P_ND$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
sort_CCF_values(amplitude.P_ND$mean, clim.daily$max.vwc)
sort_CCF_values(log10(amplitude.P_ND$mean), log10(clim.daily$max.vwc))
```

## With P_D

```{r}
cor.test(log10(amplitude.P_D$mean), log10(clim.daily$max.vwc),
  # correlation methods (pearson, spearman, kendall)
  method = c("pearson"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_D$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("spearman"),
  # set confidence interval
  conf.level = 0.95)
```

```{r}
cor.test(amplitude.P_D$mean, clim.daily$max.vwc,
  # correlation methods (pearson, spearman, kendall)
  method = c("kendall"),
  # set confidence interval
  conf.level = 0.95)
```


```{r}
ccf (amplitude.P_D$mean, clim.daily$max.vwc,
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")

ccf (log10(amplitude.P_D$mean), log10(clim.daily$max.vwc),
     #indicate what to do with "NA".
     na.action = na.pass,
     #indicate if you want the plot
     plot = "TRUE")
```

```{r}
sort_CCF_values(amplitude.P_D$mean, clim.daily$max.vwc)
sort_CCF_values(log10(amplitude.P_D$mean), log10(clim.daily$max.vwc))
```


# Seasonality ~ climate

Now, let's see if we can see a relation within Seasonality and climate.

## All trees
Let's add up all trees seasonality to a matrix with env data and explore the correlations:
```{r}
matrix <- inner_join(seasonalities.P_D.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix
```

```{r}
library("correlation")

results <- correlation(matrix)
results
```

```{r}
library(see)

results %>%
  summary(redundant = TRUE) %>%
  plot()
```
Now using spearman method:
```{r}
results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```
Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

## By class

Let's test correlations within temp and seasonality for each different class:

### Quercus Ilex
```{r}
matrix <- inner_join(seasonalities.qi.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "pearson")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

We found that after 6 instances of 15 mins (1h and half), there is 0.70 inverse correlation within surface temperature and Quercus ilex maximum swelling.

Let's plot this to visually see it for June and July Months (Daily mean seasonality + Daily temperature fluctuations)
```{r}
temp.jun_jul <- db.env %>% filter( (ts >= as.Date("2023-06-01") ) & (ts < as.Date("2023-08-01")) ) %>%  mutate(timeOfDay = as_hms(ts)) %>% summarise(meanTemp = mean(surface.temp), se_temp = (sd(surface.temp) / sqrt(n()) ), .by = timeOfDay)
temp.jun_jul

plot_day_seasonality_and_temp(seasons.qi.jun_jul, temp.jun_jul, "Quercus Ilex", PLACE, "June to July 2023")
# # ggsave(glue('output/{PLACE}_aggoneday-june-july-seasonalities&Temp-Qi.png'), width = 15, height = 10)
```

### Non-Declining Pines
```{r}
matrix <- inner_join(seasonalities.P_ND.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

Let's plot this to visually see it for June and July Months (Daily mean seasonality + Daily temperature fluctuations)
```{r}
plot_day_seasonality_and_temp(seasons.P_ND.jun_jul, temp.jun_jul, "Non-Declining Pines", PLACE, "June to July 2023")
# # ggsave(glue('output/{PLACE}_aggoneday-june-july-seasonalities&Temp-P_ND.png'), width = 15, height = 10)
```

### Declining Pines
```{r}
matrix <- inner_join(seasonalities.P_D.agg, db.env, by='ts') %>% select(soil.temp, surface.temp, air.temp, vwc, mean) %>% rename(seasonal.value = mean)
matrix

results <- correlation(matrix, method = "spearman")

results %>%
  summary(redundant = TRUE) %>%
  plot()
```

Doing cross-correlation:
```{r}
ccf(matrix$seasonal.value, matrix$surface.temp, lag.max = 102)

find_Max_CCF(matrix$seasonal.value, matrix$surface.temp)
```

Let's plot this to visually see it for June and July Months (Daily mean seasonality + Daily temperature fluctuations)
```{r}
plot_day_seasonality_and_temp(seasons.P_D.jun_jul, temp.jun_jul, "Declining Pines", PLACE, "June to July 2023")
# # ggsave(glue('output/{PLACE}_aggoneday-june-july-seasonalities&Temp-P_D.png'), width = 15, height = 10)
```

# Exploring wavelets

Here we test if there's another correlation with a longer frequency than the daily frequency between temperature and daily seasonality. We could think that, in warm periods (days, weeks, months,...) like the summer season we have a correlation with an specific seasonality.
I would expect to have a year correlation too (every summer we have similar tree), but unfortunately we don't have enough data to test this yet.

Next steps would take a long while so:
```{r}
stop('Wavelet analysis below')
```

## For all trees: surface.temp ~ seasonality

```{r}
surface.temp = db.env %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts)) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(db.env[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, surface.temp)
           
any(is.na(surface.temp))
summary(surface.temp)

seasonality = seasonalities.agg %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts), value = mean) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(seasonalities.agg[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, value)
seasonality <- data.frame(seasonality)

any(is.na(seasonality))
summary(seasonality)

# wavelet analysis (this will take a while)
wavelet = wtc(surface.temp, seasonality,
              max.scale = 32,
              #display the progress bar
              quiet = F)

# wavelet plot
par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(wavelet,
     #legend colors
     plot.cb = TRUE,
     #lag phase (lines)
     plot.phase = TRUE, 
     ylab = "Frequency (days)",
     xlab = "Time (days)") 
```

## By class: surface.temp ~ seasonality

For Quercus:
```{r}
seasonality = seasonalities.qi.agg %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts), value = mean) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(seasonalities.agg[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, value)
seasonality <- data.frame(seasonality)

any(is.na(seasonality))
summary(seasonality)

# wavelet analysis (this will take a while)
wavelet = wtc(surface.temp, seasonality,
              max.scale = 32,
              #display the progress bar
              quiet = F)

# wavelet plot
par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(wavelet,
     #legend colors
     plot.cb = TRUE,
     #lag phase (lines)
     plot.phase = TRUE, 
     ylab = "Frequency (days)",
     xlab = "Time (days)") 
```

For Non-Declining pines:
```{r}
seasonality = seasonalities.P_ND.agg %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts), value = mean) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(seasonalities.agg[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, value)
seasonality <- data.frame(seasonality)

any(is.na(seasonality))
summary(seasonality)

# wavelet analysis (this will take a while)
wavelet = wtc(surface.temp, seasonality,
              max.scale = 32,
              #display the progress bar
              quiet = F)

# wavelet plot
par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(wavelet,
     #legend colors
     plot.cb = TRUE,
     #lag phase (lines)
     plot.phase = TRUE, 
     ylab = "Frequency (days)",
     xlab = "Time (days)") 
```

For Declining pines:
```{r}
seasonality = seasonalities.P_D.agg %>% 
           # select data at hourly intervals by
           # create a new variable named "minutes"
           mutate (minutes = minute (ts), value = mean) %>% # Add variable minutes
           # filter "oclock data"
           filter (minutes == "0") %>% # Filter hourly data
           # convert datetime to a numeric variable
           # to make it more intuitive, we will express time as days,
           # and make time 0 = the first instance (corresponds with ts_start)
           mutate (time.days = (as.numeric(ts)- as.numeric(seasonalities.agg[1,"ts"]))/(60*60*24)) %>%
           # select the variables of interest (time and air.temp)
           select (time.days, value)
seasonality <- data.frame(seasonality)

any(is.na(seasonality))
summary(seasonality)

# wavelet analysis (this will take a while)
wavelet = wtc(surface.temp, seasonality,
              max.scale = 32,
              #display the progress bar
              quiet = F)

# wavelet plot
par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(wavelet,
     #legend colors
     plot.cb = TRUE,
     #lag phase (lines)
     plot.phase = TRUE, 
     ylab = "Frequency (days)",
     xlab = "Time (days)") 
```

