---
title: "Seasonality-analysis"
author: "Abel Serrano Juste"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages:
```{r}
source('lib-dendro.R')
library(glue)
library(tidyverse)
library (ggplot2)
library(lubridate)

library(imputeTS)
library(anomalize) # Decomposition
library(stats) # Decomposition
library(ggpubr) # ggarange
library(zoo)
```

Set global vars
```{r}
### DEFINE GLOBAL VARS ###
PATH = dirname(rstudioapi::getSourceEditorContext()$path)
print(PATH)
setwd(PATH)

PLACE = 'Miedes'
DATA_DIR = 'processed/Miedes-last'
ENV_DIR = 'processed/Miedes-env-processed'
```

Select dendros for this analysis
```{r}
selected_dendros <- c("92222156", "92222169", "92222175", "92222157", "92222154", "92222170", "92222173", "92222180", "92222155", "92222163", "92222171", "92222161", "92222164")
```


Load dataset:
```{r}
list_files <- list.files(file.path(".",DATA_DIR), pattern="*.csv$", full.names=TRUE)
db<-read.all.processed(list_files)
```

## CLEAN & PREPARE DATA ###

keep data of selected dendros only
```{r}
db = db %>% filter(db$series %in% selected_dendros)
str(db)
```

select from May 2023 to Sept 2023
```{r}
# Set initial and final date for analysis
ts_start <- "2023-05-01 09:00:00" # 2 days after installation
ts_end <-"2023-09-10 07:00:00" # last timestamp of downloaded data
db <- reset.initial.values(db, ts_start, ts_end)
```

# INSPECT DATA
```{r}
str(db)
head(db)
tail(db)
```

# Data imputation

Missing data
```{r}
statsNA(db$value)
```

Let's fill it through interpolation
```{r}
db$value <- db$value %>% 
            na_interpolation(option = "spline")
```

# Time Series decomposition

## Define useful functions

Define plot_decompose function which uses basic time series decomposition
```{r}
## Function decompose
plot_decompose <- function (db, name) {
  # decompose a time series (has to be in tibble format)
  decompose = time_decompose(
    # choose dataframe containing the data and convert it to tibble
    as.tibble(db,
              #what to do with na in the dataframe
               na.action = na.pass),
    # select varaible to decompose
    value)
  
  # plot time series
  p_ts = ggplot (decompose, aes(x = ts, y = observed)) +
    ggtitle(glue("Time series decomposition for {name}")) +
    geom_line (col = "black") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = NULL,
             y = "Original data (um)")
  
  # plot trend
  p_trend = ggplot (decompose, aes(x = ts, y = trend)) +
    geom_line (col = "#D55E00") + 
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    geom_hline(yintercept = 0, linetype='dotted', col = 'red') +
    theme_classic() +
    labs ( x = NULL,
             y = "Trend")
  
  # plot season
  p_season = ggplot (decompose, aes(x = ts, y = season)) +
    geom_line (col = "#E69F00") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = NULL,
             y = "Season")
  
  # plot remainder
  p_remainder = ggplot (decompose, aes(x = ts, y = remainder)) +
    geom_line (col = "#F0E442") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme_classic() +
    labs ( x = "Period (month)",
             y = "Remainder")
  
  # plot all together
  ggarrange (p_ts, p_trend, p_season, p_remainder, ncol = 1)
}

# Save the plot of decompose_ts
save_plot_decompose <- function (dat, name) {
  plot_decompose(dat, name) %>% ggexport(filename = glue('output/{name}-decomp-anomalize.png'),  width = 4500, height = 3000, res = 300)
}
```

Define amplitude function
```{r}
calculate_amplitude <- function(dat) {
  dat %>% mutate(date = date(ts)) %>% group_by(date) %>% summarize(max = max(value), min = min(value)) %>% mutate(ampl = max-min)
}

save_plot_amplitude <- function(dat.ampl, name) {
  ggplot(data = dat.ampl, mapping = aes(x = date, y = ampl)) + geom_line()
  ggsave(glue('output/{name}-amplitude.png'), width = 15, height = 10)
}
```

Define stl function
```{r}
plot_stl <- function (stl.out) {
  summary(stl.out)
  plot(stl.out)
}

save_plot_stl <- function (stl.out, name) {
  png(glue('output/{name}-stl.png'), width=15, height=10, units="in", res=300)
  plot_stl(stl.out)
  dev.off()
}

# Now plotting only seasonality but with dates in the x-axis
plot_seasonality <- function(stl.out, name) {
  
  seasonality <- stl.out$time.series[,1]
  timestamps <- seq(from = as.POSIXct(ts_start, tz='Madrid/Spain'), by = "15 min", length.out = length(dat.ts))
  zoo_data <- zoo(seasonality, order.by = timestamps)
  plot(zoo_data, xaxt = "n", type = "l", xlab = "", ylab = "Value", main = glue("Time Series by Month-Year for {name}"))
  axis(1, at = time(zoo_data), labels = format(time(zoo_data), "%Y-%m"))
  
  # Add x-axis label
  mtext("Month-Year", side = 1, line = 3)
}

save_plot_seasonality_stl <- function(stl.out, name) {
  png(glue('output/{name}-seasonality-stl.png'), width=15, height=10, units="in", res=300)
  plot_seasonality(stl.out, name)
  dev.off()
}
```

## Per individuals analysis

For each individual, calculate and plot: basic decomposition, amplitude, stl decomposition and stl seasonality

```{r}
TreeList <- read.table("TreeList.txt", header=T)

# First, define dendro number
column_names <- c("date", "min", "max", "ampl", "class")
amplitude.df = data.frame(matrix(nrow = 0, ncol = length(column_names)))
colnames(amplitude.df) <- column_names
for (dendro.no in selected_dendros) {
  
  class = (TreeList[TreeList$series == dendro.no,])$class
  name = paste(dendro.no, class, sep="-")
  
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  
  # Basic ts decomposition
  save_plot_decompose(dat, name)
  
  # Max - min daily amplitude
  dat.ampl <- calculate_amplitude(dat)
  aux <- cbind(ampl = dat.ampl, class=class) %>% rename (date = ampl.date, min = ampl.min, max = ampl.max, ampl = ampl.ampl)
  amplitude.df = rbind.data.frame(amplitude.df, aux)
  save_plot_amplitude(dat.ampl, name)
  
  # Create R TS object for stl functions
  dat.ts <- ts(data = dat$value,
             # start = dat$ts[1],
             # end = dat$ts[length(dat$ts)],
             frequency = 96)
  
  head(dat.ts, n = 180)
  
  # Decompose using STL
  stl.out = stl(dat.ts, s.window = 25, t.window = 673)
  
  # save all components
  save_plot_stl(stl.out, name)
  
  # save only seasonality
  save_plot_seasonality_stl(stl.out, name)
}
```

## Same but means of classes

Calculate mean of the seasonality for Qi, PD and P_ND
```{r}
Qi = TreeList %>% filter(class == "Quercus") %>% pull(series)
P_D = TreeList %>% filter(class == "D") %>% pull(series)
P_ND = TreeList %>% filter(class == "ND") %>% pull(series)
```


```{r}
for (dendro.no in Qi) {
  # Filter data by that no
  dat = db[db$series == dendro.no,]
  dat = dat %>% select(ts, value)
  
  stl.out = stl(dat.ts, s.window = 25, t.window = 673)
  
  list.append(stl.out)
}

mean = mean of the seasonality of every element of the list
sd = sd of list seasonalities
```


# Correlations

## Amplitude ~ climate

Mean of all amplitudes 

```{r}
amplitude.df %>% group_by(date, class) %>% summarise(mean = mean(ampl))
```




## Seasonality ~ climate

