---
title: "Seasonality-analysis"
author: "Abel Serrano Juste"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import packages:
```{r}
source('lib-dendro.R')
source('lib-ts-analysis.R')
source('correlations.R')

library(glue)
library(tidyverse)
library(ggplot2)

library(lubridate)
library(hms)

library(imputeTS)
```

Set global vars
```{r}
### DEFINE GLOBAL VARS ###
PATH = dirname(rstudioapi::getSourceEditorContext()$path)
print(PATH)
setwd(PATH)

PLACE = 'Miedes'
DATA_DIR = glue('processed/{PLACE}-processed')
ENV_DIR = glue('processed/{PLACE}-env-processed')


# Select TMS serial no for this analysis
SELECTED_TMS <- case_when(  # select one TMS sensor for this site which has all its data valid
  PLACE == 'Miedes' ~ c("94231940", "94231935"),
  PLACE == 'Corbalan' ~ c("94231943"),
  PLACE == 'Penaflor' ~ c("94231934"), # 94231950 is delayed? and throws wrong results? double check
  .default = NA
)

TreeList <- read.table("TreeList.txt", header=T)

ALL_PERIOD_ST <- "May to September (all study period)"
```

Select dendros for this analysis (all if FALSE)
```{r}
if (PLACE == 'Miedes') {
  selected_dendros <- c("92222156", "92222169", "92222175", "92222157", "92222154", "92222170", "92222173", "92222180", "92222155", "92222163", "92222171", "92222161", "92222164")  
} else {
 selected_dendros <- c() 
}
```

Load datasets:
```{r}
list_files <- list.files(file.path(".",DATA_DIR), pattern="*.csv$", full.names=TRUE)
db<-read.all.processed(list_files)

db.env <- read.env.proc(file.path(".",ENV_DIR,'proc-env.csv'))
```
# CLEAN & PREPARE DATA #

keep data of selected dendros only
```{r}
if (length(selected_dendros) > 0) {db = db %>% filter(db$series %in% selected_dendros)}
str(db)
```

select from May 2023 to Sept 2023
```{r}
# Set initial and final date for analysis
ts_start <- "2023-05-01 09:00:00" # from first of May
ts_end <-"2023-09-10 07:00:00" # to 10th of september
db <- reset.initial.values(db, ts_start, ts_end)
db.env <- db.env[which(db.env$ts>=ts_start & db.env$ts<=ts_end),]
```

Add classes column:
Define set of series for each class
```{r}
Qi = TreeList %>% filter(class == "Quercus", series %in% unique(db$series)) %>% pull(series)
P_D = TreeList %>% filter(class == "D", series %in% unique(db$series)) %>% pull(series)
P_ND = TreeList %>% filter(class == "ND", series %in% unique(db$series)) %>% pull(series)
```


```{r}
db <- db %>%  mutate (class = case_when(
                    series %in% Qi ~ factor("Quercus"),
                    series %in% P_D ~ factor("D"),
                    series %in% P_ND ~ factor("ND"),
                    .default = NA
                    )
              )
```

Calculate means of selected TMSs only:
```{r}
db.env <- db.env %>% filter(series %in% SELECTED_TMS)

# interpolate NAs with:
if (any(is.na(db.env$vwc))) {
  for (series_no in SELECTED_TMS) {
    db.env[db.env$series == series_no,] <- db.env[db.env$series == series_no,] %>%
      mutate_at(vars(air.temp, surface.temp, soil.temp, vwc), \(x) (na_interpolation(x, option = "spline")))
  }
}

statsNA(db.env$vwc)

db.env <- db.env %>% group_by(ts) %>% summarise(soil.temp = mean(soil.temp), surface.temp = mean(surface.temp), air.temp = mean(air.temp), vwc = mean(vwc)) 
```

Plot env data:
```{r}
plot_temp_and_humidity <-
  ggplot(data = db.env, aes(x=ts, y=surface.temp)) +
    ggtitle(glue("Temperature and humidity for {PLACE} (May to September 2023)")) +
    geom_line(aes(colour = surface.temp)) +
    theme_bw() +
    scale_colour_gradient(low = "light blue", high = "red") +
    scale_x_datetime(date_breaks = "1 month", date_labels="%b %Y") +
    theme(axis.text.x = element_text(angle = 30, hjust=1)) +
    scale_y_continuous("Temperature (ºC)", breaks=seq(0,50,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
    geom_line(aes(x = ts, y = vwc * 100,  linetype = "Volumetric Water Content (%)"), col="blue", show.legend	= TRUE) +
    scale_linetype_manual(NULL, values = 1) +
    labs(x=expression(''), colour = "Temperature (ºC)")
plot_temp_and_humidity
# ggsave(glue('output/{PLACE}-VWCytempAgg-selected-period.png'), width = 15, height = 10)
```

Set ggplot label options, themes & vars:

```{r}
theme_set( theme_bw() +
          theme(
            axis.text=element_text(size=12),
            legend.text=element_text(size=10),
            axis.title=element_text(size=14),
            plot.title = element_text(hjust = 0.5, face = "bold")
          ))

every_hour_labels = format(lubridate::parse_date_time(hms(hours = 0:23), c('HMS', 'HM')), '%H:%M')
labels = c("Quercus", "Declining Pines", "Non-Declining Pines")

n.per.class <- db %>% group_by(class) %>% summarise (n = n_distinct(series)) %>% pivot_wider(names_from = class, values_from = n)

c_labels = c(glue("{labels[1]} ({n.per.class$Quercus})"), glue("{labels[2]} ({n.per.class$D})"), glue("{labels[3]} ({n.per.class$ND})"))
c_values = c(Quercus = "purple", D = "darkred", ND = "darkorange")
scale_colour_custom <- function(...) scale_color_manual(labels = c_labels, values=c_values)
options(ggplot2.discrete.colour = scale_colour_custom)
```


### Data imputation for dendros

Missing data
```{r}
statsNA(db$value)
```

Let's fill it through interpolation
```{r}
db$value <- db$value %>% 
            na_interpolation(option = "spline")
```

### Inspect data

#### dendros
```{r}
str(db)
head(db)
tail(db)
```

#### climate
```{r}
head(db.env)
summary(db.env)
```

### Others
Calculations of useful aggregation data for climate:
  * soil moisture: max VWC, range VWC, mean of VWC (misleading)
  * temp: Mean of daily temp, minimum daily temp and maximum daily temp, range temp and interquartile temp (Q3-Q1). We will use soil temperature as it's used on many other studies and it isn't so much influenced by sun irradiation and night/day differences.

```{r}
clim.daily <- db.env %>% mutate (date = date(ts)) %>% group_by(date) %>% summarise(max.temp = max(surface.temp), min.temp = min(surface.temp), mean.temp = mean(surface.temp), sd.temp = sd(surface.temp), range.vwc = max(vwc) - min(vwc), max.vwc = max(vwc), mean.vwc = mean(vwc), range.temp = max.temp - min.temp, interquartil.temp = as.numeric(quantile(surface.temp, prob=c(.75)) - quantile(surface.temp, prob=c(.25)) ) )
summary(clim.daily)

clim.hourly <- db.env %>% mutate (date = date(ts), hour = hour(ts)) %>% group_by(date, hour) %>% summarise(max.temp = max(surface.temp), min.temp = min(surface.temp), mean.temp = mean(surface.temp), sd.temp = sd(surface.temp), range.vwc = max(vwc) - min(vwc), max.vwc = max(vwc), mean.vwc = mean(vwc), range.temp = max.temp - min.temp, interquartil.temp = as.numeric(quantile(surface.temp, prob=c(.75)) - quantile(surface.temp, prob=c(.25))), .groups = "drop" )
summary(clim.hourly)
```

# TIME SERIES DECOMPOSITION

Extract trend, residuals and seasonality for all trees and by class:

```{r}
decomposed.db <- decompose_ts_stl(db, selected_dendros)
decomposed.db <- decomposed.db %>% mutate (
                date = date(ts),
                hour = hour(ts),
                class = case_when(
                    series %in% Qi ~ factor("Quercus"),
                    series %in% P_D ~ factor("D"),
                    series %in% P_ND ~ factor("ND"),
                    .default = NA
                    )
                )
decomposed.db
```

Time-series decomposition by class
```{r}
decomposed.db.alltrees <- decomposed.db %>% group_by(ts) %>% summarise(seasonal = mean(seasonal), trend = mean(trend), remainder = mean(remainder), .groups = "drop")
decomposed.db.byclass <- decomposed.db %>% group_by(ts, class) %>% summarise(seasonal = mean(seasonal), trend = mean(trend), remainder = mean(remainder), .groups = "drop")
```

Calculate time-series decomposition in aggregated dataframes (daily)

```{r}
decomposed.db.daily.alltrees <- decomposed.db %>% summarise(seasonal = mean(seasonal), trend = mean(trend), remainder = mean(remainder), .by = date)
decomposed.db.daily.byclass <- decomposed.db %>% summarise( seasonal = mean(seasonal), trend = mean(trend), remainder = mean(remainder), .by = c(date, class))
```

Calculate growth rate: !!! Cambiar por growth rate ratio mejor??

```{r}
db.gr <- data.frame()
for (dendro.no in selected_dendros) {
  # Filter data by that no
  dat = decomposed.db[decomposed.db$series == dendro.no,] %>% select(ts, trend, class) %>% rename(value = trend)
  
  aux <- data.frame(
     series = as.factor(dendro.no),
     class = dat$class[1],
     calc.growth.rate(dat)
  ) 
  db.gr <- rbind.data.frame(db.gr, aux)
}

db.gr

db.gr.alltrees <- db.gr %>% summarise( mean_gr = mean(rate, na.rm = T), gr_se = sd(rate / sqrt(n()), na.rm = T), .by = c(date)) %>% mutate(doy = yday(date))
db.gr.byclass <- db.gr %>% summarise( mean_gr = mean(rate, na.rm = T), gr_se = sd(rate / sqrt(n()), na.rm = T), .by = c(date, class)) %>% mutate(doy = yday(date))
```

# Determining the time frequency for each comparison:

residuals ~ climate -> every 15': it should be each 15' or each hour, since the remainder is what excess from the seasonality, which we treated each 15'. 
trend ~ climate -> daily: Better daily, as the trend takes some time to change and it makes sense also for climate variables: max.vwc stays some days after a rain episode.
growth rate ~ climate -> daily: This has to be treated daily since the growth rate is a daily rate.
TWD ~ climate -> daily. acumulado por momentos diarios

# Plots

plot trend ~ vwc
```{r}
ggplot(data = decomposed.db.daily.byclass, mapping = aes(x = date, y = trend, col = class)) +
  ggtitle('Compare trend vs soil moisture') +
  geom_line() +
  scale_y_continuous("Trend (um per day)", breaks=seq(0,900,100), sec.axis = sec_axis(trans = ~ . /1000, name = "Volumetric Water Content (%)")) +
  geom_line(data = clim.daily, mapping = aes(x = date, y = max.vwc * 1000, linetype = "Max Volumetric Water Content (%)"), col="blue")
```

plot trend ~ temperatures
```{r}
ggplot(data = decomposed.db.daily.byclass, mapping = aes(x = date, y = trend, col = class)) +
  ggtitle('Compare trend vs temperature') +
  geom_line() +
  scale_y_continuous("Trend (um per day)", breaks=seq(0,900,100), sec.axis = sec_axis(trans = ~ . /20, name = "Temperature (ºC)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = range.temp*20, linetype = "Range Temp (ºC)"), col="firebrick1")
```

plot residuals ~ vwc

```{r}
ggplot(data = decomposed.db.byclass, mapping = aes(x = ts, y = remainder, col = class)) +
  ggtitle('Compare residuals vs soil moisture') +
  geom_line() +
  scale_y_continuous("Remainder (um per day)", sec.axis = sec_axis(trans = ~ . /100, name = "Volumetric Water Content (%)")) +
  geom_line(data = db.env, mapping = aes(x = ts, y = vwc * 100, linetype = "Max Volumetric Water Content (%)"), col="blue")
```


plot residuals ~ temperatures
```{r}
ggplot(data = decomposed.db.byclass, mapping = aes(x = ts, y = remainder, col = class)) +
  ggtitle('Compare residuals vs Temperature') +
  geom_line() +
  scale_y_continuous("Remainder (um per day)", sec.axis = sec_axis(trans = ~ . /2, name = "Temperature (ºC)" ))+
  geom_line(data = db.env, mapping = aes(x = ts, y = surface.temp*2, linetype = "Range Temp (ºC)"), col="firebrick1")
```

plot growth rate ~ vwc

```{r}
ggplot(data = db.gr.byclass, mapping = aes(x = date, y = mean_gr, col = class)) +
  ggtitle('Compare soil moisture with daily growth rate') +
  scale_color_manual(labels = c_labels, values=c_values) +
  geom_line() +
  scale_y_continuous("Growth rate (um per day)", breaks=seq(0,100,10), sec.axis = dup_axis(name = "Volumetric Water Content (%)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = max.vwc * 100, linetype = "Max Volumetric Water Content (%)"), col="blue")
```

plot growth rate ~ temperatures

```{r}
ggplot(data = db.gr.byclass, mapping = aes(x = date, y = mean_gr, col = class)) +
  ggtitle('Compare temperature with daily growth rate') +
  geom_line() +
  scale_y_continuous("Growth rate (um per day)", breaks=seq(0,100,10), sec.axis = dup_axis(name = "Temperature (ºC)" ))+
  geom_line(data = clim.daily, mapping = aes(x = date, y = range.temp, linetype = "Range Temp (ºC)"), col="firebrick1")
```


# Correlations

## trend ~ climate

All trees:
- Soil moisture

```{r}
cor.test(decomposed.db.daily.alltrees$trend, clim.daily$max.vwc)
```
^ Nope

```{r}
ccf (decomposed.db.daily.alltrees$trend, clim.daily$max.vwc, lag.max = 50, plot = "TRUE")
sort_CCF_values(decomposed.db.daily.alltrees$trend, clim.daily$max.vwc, lag.max = 50)
```

All trees ~ temp
```{r}
ccf (decomposed.db.daily.alltrees$trend, clim.daily$range.temp, lag.max = 50, plot = "TRUE")
sort_CCF_values(decomposed.db.daily.alltrees$trend, clim.daily$range.temp, lag.max = 50)
```

Quercus:

```{r}
ccf (decomposed.Qi$trend, clim.hourly$max.vwc, lag.max = 168, plot = "TRUE")
sort_CCF_values(decomposed.Qi$trend, clim.hourly$max.vwc, lag.max = 168)
```

```{r}
clim.hourly
```


Non-Declining Pines:

Declining pines:


## residuals ~ climate
All trees ~ moisture
```{r}
ccf (decomposed.db.alltrees$remainder, db.env$vwc, lag.max = 96, plot = "TRUE")
sort_CCF_values(decomposed.db.alltrees$remainder, db.env$vwc, lag.max = 96)
```
^ Some correlation 10 hours after a rain episode.

Quercus ~ Soil moisture
```{r}
Qi <- decomposed.db.byclass[decomposed.db.byclass$class == 'Quercus',]
ccf (Qi$remainder, db.env$vwc, lag.max = 96, plot = "TRUE")
sort_CCF_values(Qi$remainder, db.env$vwc, lag.max = 96)
```
^ Some correlation ~2 hours after a rain episode.


All trees ~ temp
```{r}
ccf (decomposed.db.alltrees$remainder, db.env$surface.temp, lag.max = 96, plot = "TRUE")
sort_CCF_values(decomposed.db.alltrees$remainder, db.env$surface.temp, lag.max = 96)
```
^ There's no good correlation

## growth rate ~ climate

All trees ~ moisture
```{r}
ccf (db.gr.alltrees$mean_gr, clim.daily$max.vwc, lag.max = 30, plot = "TRUE")
sort_CCF_values(db.gr.alltrees$mean_gr, clim.daily$max.vwc, lag.max = 30)
```

All trees ~ temp
```{r}
cor.test(db.gr.alltrees$mean_gr, clim.daily$range.temp)
cor.test(db.gr.alltrees$mean_gr, clim.daily$mean.temp)
cor.test(db.gr.alltrees$mean_gr, clim.daily$max.temp)
cor.test(db.gr.alltrees$mean_gr, clim.daily$min.temp)
```

```{r}
ccf (db.gr.alltrees$mean_gr, clim.daily$range.temp, lag.max = 30, plot = "TRUE")
sort_CCF_values(db.gr.alltrees$mean_gr, clim.daily$range.temp, lag.max = 30)
```

